---
output: html_document
editor_options: 
  chunk_output_type: console
---


<pre>
## Callin Switzer
## Analysis of bee intrumental learning hypothesis experiment
### i.e. Can bees learn to sonicate at low or high frequencies?

## Initial commit: 17 Nov 2017
## Updated 15 Dec, cleaned code for submission
## Update 14 Jan 2019 prepped for depositing in zenodo
</pre>


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
```

```{r, warning = FALSE, message=FALSE}
#install packages
ipak <- function(pkg){
     new.pkg <- pkg[!(pkg %in% installed.packages()[, "Package"])]
     if(length(new.pkg)) install.packages(new.pkg, dependencies = TRUE)
     sapply(pkg, require, character.only = TRUE)
}

packages <- c("plyr", "tidyverse", 'lme4', 'sjPlot', 
              "multcomp", "cowplot")
ipak(packages)

# set ggplot theme
theme_set(theme_classic() + 
            theme(axis.text=element_text(colour="black"), 
                  text=element_text(size=10)))

# set  directories
dataDir <- file.path(getwd(), "data")
figDir <- file.path(getwd(), "figures")

print(paste("last run ", Sys.time()))
print(R.version)
```

# Read in data and tidy

```{r, message=FALSE}
# read in data
sl <- read_csv(file.path(dataDir, '01_CombinedTrials_cleaned.csv'))

# clean data
sl <- sl %>%
  # gets rid of the copious warnings
  mutate(timeSinceStart = NA, 
         timeSinceLastBuzz = NA, 
         buzzesSinceReward = NA) %>%
  
  # make sure all bee colors are lowercase
  mutate(beeCol =  tolower(beeCol)) %>%
  
  # fix orange.5, which should be orange.3
  mutate(hive = ifelse(hive == "5" & beeCol == "orange", "3", hive)) %>%
  
  # color + hive is an ID var 
  mutate(beeColHive = interaction(beeCol, hive), 
         colNum = paste(beeCol, hive, sep = "_")) %>%
  
  # fix IT for orange.3
  mutate(IT = replace(IT, beeColHive == "orange.3", 3.99)) %>%
  
  # remove two trials that contain errors
  filter(!(beeCol == "whitepink" & trialNum == 3)) %>%
  filter(!(beeCol == "limepurple" & trialNum == 3)) %>%
  

  # convert hive to factor
  mutate(hive = as.factor(hive),
         trt2  = ifelse(trt == 'full' & trialNum > 1, 
                        "full_2", as.character(trt)), 
         trt = relevel(factor(trt), ref = "full")) %>%
  
  # divide amplitud by 2
  mutate(amp_acc2 = amp_acc / 2) %>%
  mutate(trialNum0 = trialNum - 1)



# calculate mean IT span (for centering)
IT_mean <- sl %>% 
    group_by(beeColHive) %>% 
    slice(1) %>%
  ungroup() %>%
  dplyr::select(IT_imputed) %>%
  summarise(mean = mean(IT_imputed)) %>%
  as.numeric()

# print overall mean of IT span
print(IT_mean)

# add new variable 
# that is the centered IT span
sl <- sl %>% 
  mutate(IT_centered = IT_imputed - IT_mean)

```


# Calculate more descriptive stats 
# and make supplementary fig for paper

```{r}
# std dev for rewards by treatment
s2 = sl %>%
  mutate(beeColHive = interaction(beeCol, hive), 
         reward= grepl(pattern = "t", x = .$rewTF, ignore.case = TRUE)*1) 


s3 = s2 %>%
  group_by(beeColHive, trt2) %>%
  dplyr::summarize(mean_reward = mean(reward)) %>%
  group_by(trt2) %>%
  dplyr::arrange(desc(trt2)) 

# summary for paper
s3 %>%
  dplyr::summarize( mean_ = mean(mean_reward), 
                    reward_sd = sd(mean_reward), n = n()) %>%
  mutate(se = reward_sd / sqrt(n))


# total number of trials for each treatment
tapply((grepl("[tT]", as.character(sl$rewTF))), INDEX= sl$trt, length)

# total number of rewards per treatment
tapply((grepl("[tT]", as.character(sl$rewTF))), 
       INDEX= sl$trt, FUN = function(x) sum(x))

# total number of trials that were unrewarded per treatment
tapply((grepl("[tT]", as.character(sl$rewTF))), 
       INDEX= sl$trt, FUN = function(x) sum(!x))


# calculate avg freq & SE
se <- function(x) {
  sd(x) /  sqrt(length(x))
}


s4 = s2 %>%
  group_by(beeColHive, trt2) %>%
  dplyr::summarize(mean_freq = mean(freq), 
                   sd_freq = sd(freq)) %>%
  group_by(trt2) %>%
  dplyr::arrange(desc(trt2)) %>%
  filter(trt2 != "full") %>%
  ungroup() %>%
  mutate(trt3 = plyr::mapvalues(.$trt2, from = c("full_2", "high", "low"), 
                        to = c("Full range\n(220 - 450 Hz)", 
                               "High range\n(340 - 390 Hz)", 
                               "Low range\n(220 - 330 Hz)")))


# plot bee size in the groups
sl_sum <- sl %>%
  dplyr::select(beeColHive, trt2, IT_imputed) %>%
  distinct(beeColHive, trt2, .keep_all = TRUE) 

sl_sum %>%
  group_by(trt2) %>%
  summarize(counts = n())

sl %>% nrow

sl %>% group_by(beeColHive) %>%
  summarize(n()) %>%
  nrow()

sl %>% group_by(trt2) %>%
  summarize(count= n())

s2 <- sl_sum  %>%
  filter(trt2 != "full") %>%
  mutate(trt3 = plyr::mapvalues(.$trt2, from = c("full_2", "high", "low"), 
                        to = c("Full range\n(220 - 450 Hz)", 
                               "High range\n(340 - 390 Hz)", 
                               "Low range\n(220 - 330 Hz)")))
# join data  for plotting
s44 <- full_join(s2, s4)

set.seed(123)
ggplot(s44, aes(x = trt3,  y= IT_imputed)) + 
  geom_violin(width = 0.5, fill = "grey90", color = NA, bw = 0.2, trim = TRUE) + 
  #geom_boxplot(width = 0.5, outlier.alpha =   0) + 
  geom_point(position = position_jitter(height = 0, width = 0.05), 
             stroke= 0, size = 2.5, color = 'grey40') + 
  labs(x = "Frequency range for reward", y = "Intertegular span (mm)") 


# supplementary figure for paper
ggsave(file.path(figDir, "beeSize_exp1.tiff"), 
       width = 6, height = 4, units = "in", dpi = 500)
```

# Plot frequency vs. acceleration

```{r}
# plot freq vs. amp
sl <- sl %>%
  mutate(`IT interval (mm)` = cut_interval(IT_imputed, n = 3))

ggp <- ggplot(sl[sl$trt2 == "full", ], aes(x = freq, y = amp_acc2)) + 
  geom_point(position = position_jitter(height = 0, width = 2), 
             alpha = 0.3, stroke = 0, size = 1.5) + 
  stat_smooth(method = 'loess', span = 1, color = 'grey40', se = FALSE, 
              lwd = 1.3) + 
  facet_wrap(~`IT interval (mm)`, labeller = "label_both") + 
  theme(strip.background = element_blank(), 
        strip.text = element_text(size = 10)) + 
  theme(legend.position = "none", 
        plot.margin = unit(c(1,1,1,1)/2, "cm")) + 
  labs(x = "Sonication frequency (Hz)", 
       y = expression ("Sonication acceleration "(m~s^{-2})))

ggp

# supplementary figure for paper
ggsave(file.path(figDir, "freq_amp.tiff"), width = 6.5, 
       height = 3, units = "in", dpi = 500)
```


# Statistical models for frequency

```{r}
# start with large model
# fixed effects: treatment (rewarded at high, low, or full range of frequencies), 
# IT span, colony, trial number, and the interaction of treatment and IT span.
# We included a random intercept of bee ID and added a random slope of trial number 
# within bee ID. We allowed for correlation between bee ID and trial number. 

mFull = lmer(freq ~ trt + IT_centered + hive + trialNum + 
            trt:IT_centered + (1+trialNum|beeColHive), 
          data = sl, REML = FALSE)
summary(mFull)

m1 = update(mFull, .~. - IT_centered:trt)
BIC(m1, mFull) # keep m1 (without IT_imputed:trt)
summary(m1)$coef

m2 <- update(m1, .~. - hive)
BIC(m1, m2) # keep m2 (without hive)
summary(m2)$coef

m3 <- update(m2, .~. - IT_centered)
BIC(m2, m3) # keep m2 (leave IT in model)

m4 <- update(m2, .~. - trt)
BIC(m2, m4) # keep (m2) trt


# refit best model with REML = TRUE
m_Final <- update(m2, .~., REML = TRUE)
summary(m_Final) # summary for paper

# show full model for paper
fullMod_freq <- update(mFull, .~., REML = TRUE)
summary(fullMod_freq)


# m_Final diagnostics
plot(m_Final)
qqnorm(ranef(m_Final)$beeColHive[[1]])
qqline(ranef(m_Final)$beeColHive[[1]])

## bonf adjusted pvals for paper
summary(glht(m_Final, linfct = mcp(trt = "Tukey")), 
        test = adjusted("bonf"))


# ________________________________________
# refit model for bootstrap,
# without centered IT Span
# ________________________________________

m3 <- lmer(freq ~ trt + IT_imputed + trialNum +
                   (1 + trialNum|beeColHive), 
                REML = TRUE, 
                data = sl)

```

### Predict some values with frequency model

```{r}
# generate data frame for predictions
pframe0 <- data.frame(expand.grid(trt = levels(sl$trt), 
                                 IT_centered = c(-1, 
                                                  0, 
                                                1),  
                                 beeColHive  = 99999, 
                                 trialNum = 2, 
                                 freq = 0))

pframe0$preds <- predict(m_Final, newdata = pframe0, re.form=NA, type = 'response')

pframe0 %>%
  filter(IT_centered == 0)



pframe0 %>%
  filter(trt == "full")
```


# Generate Bootstrap CI's for frequency
```{r bootstrap, cache = TRUE}
# set number of bootstrap samples
nbootSims = 10000

# generate data frame for predictions
pframe <- data.frame(expand.grid(trt = levels(sl$trt), 
                                 IT_imputed = seq(min(s2$IT_imputed), 
                                                  max(s2$IT_imputed), 
                                                  length.out = 50),  
                                 beeColHive  = 99999, 
                                 trialNum = 2, 
                                 freq = 0))

pp <- predict(m3, newdata = pframe, re.form=NA, type = 'response') 


### Calculate CI's (using bootstrap, not accounting for random effects)
system.time({
  bb2 <- bootMer(m3, FUN=function(x){predict(x, 
                                             pframe, 
                                             re.form=NA, 
                                             type = 'response')}, 
                 nsim = nbootSims)
})

print(paste("Number of bootstrap samples", nrow(bb2$t)))

# calculate quantiles
bb2_se <-apply(bb2$t,2,function(x) quantile(x, probs = c(0.025, 0.975)))
pframe$blo<-bb2_se[1,]
pframe$bhi<-bb2_se[2,]
pframe$predMean <- pp
pframe <- pframe[, c('trt',"IT_imputed", "blo", "bhi", "predMean")]
```

# Make frequency plot for paper

```{r}
# rename labels for plot
pframe$trt3 = plyr::mapvalues(pframe$trt, from = c("full", "high", "low"), 
                        to = c("Full range\n(220 - 450 Hz)", 
                               "High range\n(340 - 390 Hz)",
                               "Low range\n(220 - 330 Hz)"))
set.seed(1234)
g1 <- ggplot(pframe, aes(x=IT_imputed, y=predMean, color = trt3, fill = trt3))+
        labs(y = "Sonication frequency (Hz)", x = "Intertegular span (mm)") + 
        geom_ribbon(aes(ymin = blo, ymax = bhi), alpha = 0.2, color = NA)+

        geom_line()+
        theme(plot.background = element_rect(fill = "transparent",colour = NA), 
            panel.background = element_rect(fill = "transparent",colour = NA), 
            legend.position =c(0.8, 0.8), 
            legend.background = element_rect(fill="transparent", colour = NA),
            strip.background = element_blank(),
            legend.title = element_text(size = 8),
            panel.spacing.x = unit(1, "lines"), 
           plot.margin = margin(c(0.2,0.51, 0.2, 0.2), unit = "cm")) +
        #scale_x_continuous(breaks = seq(3.4, 4.8, by = 0.5)) +
        scale_color_viridis_d(name = "Frequency range\nfor reward", 
                              end = 0.7, option = "inferno") + 
        scale_fill_viridis_d(name = "Frequency range\nfor reward", 
                             end =0.7, option = 'inferno')
g1
```


# Visualization of frequency over time

```{r, cache = TRUE}
st <- sl %>%
# gets rid of the copious warnings
  mutate(timeSinceStart = NA, timeSinceLastBuzz = NA, buzzesSinceReward = NA) %>%
  mutate(beeColHive = interaction(beeCol, hive)) %>%
  # remove two trials I messed up
  filter(!(beeCol == "whitepink" & trialNum == 3)) %>%
  filter(!(beeCol == "limepurple" & trialNum == 3)) %>%
  mutate(t5 = recode(.$trt, "full" = "Full range\n(220 - 450 Hz)", 
                       "high" = "High range\n(340 - 390 Hz)\nsecond trial", 
                       "low" = "Low range\n(220 - 330 Hz)\nsecond trial")) %>%
  mutate(t6 = t5, 
         t6 = ifelse(.$trt == 'full' & .$trialNum == 1, 
                     "Full range \n(220 - 450 Hz)\nfirst trial", t5), 
         t6 = ifelse(.$trt == 'full' & .$trialNum > 1, 
                     "Full range \n(220 - 450 Hz)\nsecond trial", t6)
         ) %>%
  filter(trialNum <= 10 & trialNum > 1) 

fta <- sl %>% 
  filter(trialNum == 1) %>% 
  dplyr::group_by(beeColHive) %>%
  dplyr::summarise(initialFreq = mean(freq)) %>%
  full_join(sl) %>%
  mutate(freqDiff = freq- initialFreq) %>%
  filter(beeColHive %in% .$beeColHive[.$trialNum > 1]) %>%
  dplyr::arrange(trt) 

cc <- fta %>%
  ungroup() %>%
  filter(trialNum > 1) %>%
  group_by(beeColHive) %>%
  dplyr::summarize(meanFreq = mean(freq))



bb <- fta %>%
  dplyr::select(beeColHive, trt, trialNum, initialFreq) %>%
  dplyr::filter(trialNum == 2 | trialNum == 3) %>%
  group_by(beeColHive) %>%
  slice(1:1) %>%
  full_join(cc) %>%
  mutate(freqDiff = initialFreq - meanFreq) %>%
  dplyr::arrange(trt, desc(freqDiff))

fta <- fta %>%
  mutate(beeColHive = factor(beeColHive, levels = bb$beeColHive)) %>% 
  mutate(t5 = recode(.$trt, "full" = "Full range\n(220 - 450 Hz)", 
                       "high" = "High range\n(340 - 390 Hz)", 
                       "low" = "Low range\n(220 - 330 Hz)")) %>%
  filter(trialNum <= 10)

s44 = fta %>%
  group_by(trialNum, beeColHive, t5) %>%
  dplyr::summarise(meanFreq2 = mean(freq))
s44

# plot to show trials
ggplot(fta, aes(x = trialNum, y = freq, color = t5)) + 
  geom_ribbon(data = fta[fta$trialNum > 1,], 
              aes(x = rep(seq(1.5, 12, 3),  
                          length.out = nrow(fta[fta$trialNum > 1,])), 
                  ymin = lowFrq, ymax = highFrq), alpha = 0.2, color = NA) + 
  geom_point(position = position_jitter(width = 0, height = 10), 
             alpha= 0.2, stroke = 0, aes( shape = t5 )) + 
  facet_wrap(~beeColHive, ncol = 8) +
  geom_line(data = s44, aes(x = trialNum, y = meanFreq2), color = 'black', 
            lwd = 0.8, show.legend = FALSE) + 
  geom_line(data = s44, aes(x = trialNum, y = meanFreq2), color = 'black', 
            lwd = 0.8, show.legend = FALSE) + 
  geom_point(data = s44, aes(x = trialNum, y = meanFreq2), 
             color = 'black', pch = 18, show.legend = FALSE) + 
  geom_ribbon(data = fta[fta$trialNum > 1,], 
              aes(x = rep(seq(0.5, 1.5, 1),  
                          length.out = nrow(fta[fta$trialNum > 1,])), 
                  ymin = 220, ymax = 450), alpha = 0.2, color = NA) + 
  scale_x_continuous(breaks = seq(0, 10, 2)) + 
  scale_color_viridis_d(name = "",
                        option = "magma", begin = 0.3,
                        end = 0.80, 
                        guide = guide_legend(override.aes = 
                                               list(alpha = 1, size = 4))) + 
  theme_classic() + 
  theme(strip.background = element_blank(), 
        strip.text = element_blank(), 
        panel.border = element_rect(colour = "grey40", 
                                    fill=NA, size=0.5), 
        legend.position = 'top') + 
  scale_shape_manual(name = "", values = c(16,17,15)) + 
   
  labs(y = "Frequency (Hz)", x = "Trial Number")

# supplementary figure for paper
ggsave(filename = file.path(figDir,"freqLearningCurve.tiff"),
       dpi = 500,  width = 6.5*1.5, height = 4*1.5, units = "in")
```

# Compare frequency analysis when 2 bees without IT are missing
```{r}
summary(m3)$coef
missing_m3 <- update(m3, .~. - IT, data = sl[!is.na(sl$IT),])

summary(missing_m3)$coef
```

___

# Statistical analysis for acceleration

```{r ampAnalysis}
# show heteroskedasticity
maa0 = lmer(amp_acc2 ~ trt* IT_centered + hive + trialNum  
            + (1+trialNum|beeColHive), data = sl, REML = FALSE)
plot(maa0)


# note log-transformation to make model fit assumptions better
maa0 = lmer(log(amp_acc2) ~ trt* IT_centered + hive + trialNum  
            + (1+trialNum|beeColHive), data = sl, REML = FALSE)
plot(maa0)

# main effect model
maa1 = update(maa0, .~. - trt:IT_centered)
BIC(maa0, maa1) # use no interaction (keep maa1)

maa2 = update(maa1, .~. - trt)
BIC(maa1, maa2) # keep treatment (maa1)

maa3 <- update(maa1, .~. - hive)
BIC(maa1, maa3) # remove hive (maa3)

maa4 <- update(maa3, .~. - IT_centered)
BIC(maa3, maa4) # remove IT, according to BIC

# we will keep IT, because it was found in the 
# frequency model, and we want our models to be comparable

# summary for paper
m13 <- update(maa3, .~., REML = TRUE)
summary(m13)



# full model summary for paper
full_acc_model <- update(maa0, .~., REML = TRUE)
summary(full_acc_model)
round(summary(full_acc_model)$coef, digits = 3)

# diagnostics
plot(m13)
qqnorm(ranef(m13)$beeColHive[[1]])
qqline(ranef(m13)$beeColHive[[1]])

# post-hoc tests with bonf adjustment
summary(glht(m13, linfct = mcp(trt = "Tukey")), 
        test = adjusted("bonf"))


# refit model on orignal scale, for better plotting
maa3 <- lmer(log(amp_acc2) ~ trt + IT_imputed + trialNum + 
              (1 + trialNum | beeColHive), 
            data = sl,  REML = TRUE)
summary(maa3)

```

### predict some values with acceleration model

```{r}

summary(m13)$coef
# generate data frame for predictions
pframe1 <- data.frame(expand.grid(trt = levels(sl$trt), 
                                 IT_centered = c(-1, 
                                                  0, 
                                                1),  
                                 beeColHive  = 99999, 
                                 trialNum = 2, 
                                 freq = 0))

pframe1$preds <- predict(m13, newdata = pframe1, re.form=NA, type = 'response')


# for treatment high
# exp(0.182265402) =  1.199933
pframe1 %>%
  filter(IT_centered == 0) %>%
  mutate(exp_preds = exp(preds))
1.199933 * 21.10552


pframe1 %>%
  filter(IT_centered == 1) %>%
  mutate(exp_preds = exp(preds))
1.199933 * 28.76815


# for IT span 
# > exp(0.309734388)
# [1] 1.363063
pframe1 %>%
  filter(trt == "full") %>%
  mutate(exp_preds = exp(preds))
1.36 * 21.105
1.36 * 15.48389

```

# Make acceleration plot the same as the frequency plot

```{r, cache = TRUE}
# don't need hive, because that's not in the model we chose (above)
pframe_a <- data.frame(expand.grid(trt = levels(sl$trt), 
                                 IT_imputed = seq(min(s2$IT_imputed), max(s2$IT_imputed), 
                                                  length.out = 50),  
                                 beeColHive  = 99999, 
                                 trialNum = 2))
pframe_a$acc <- 0

# exponentiate to put back onto original scale
pp <- predict(maa3, 
              newdata = pframe_a, 
              re.form=NA, 
              type = 'response') %>% exp() 




### Calculate CI's (using bootstrap, not accounting for random effects)
system.time({
  bb2 <- bootMer(maa3, 
                 FUN=function(x) predict(x, 
                                         pframe_a, 
                                         re.form=NA, 
                                         type = 'response'), 
                 nsim = nbootSims) 
})

print(paste("Number of bootstrap samples", nrow(bb2$t)))
bb2_se <-apply(bb2$t,2,function(x) quantile(x, probs = c(0.025, 0.975)))
pframe_a$blo<-bb2_se[1,] %>% exp()
pframe_a$bhi<-bb2_se[2,] %>% exp()
pframe_a$predMean <- pp
pframe_a <- pframe_a[, c('trt',"IT_imputed", "blo", "bhi", "predMean")]
```


# Make accel plot for paper

```{r}
# recode factor
pframe_a$trt3 = plyr::mapvalues(pframe$trt, from = c("full", "high", "low"), 
                        to = c("Full range\n(220 - 450 Hz)", 
                               "High range\n(340 - 390 Hz)",
                               "Low range\n(220 - 330 Hz)"))
ga1 <- ggplot(pframe_a, aes(x=IT_imputed, y=predMean, color = trt3, fill = trt3))+
        
        labs(y = expression ("Sonication acceleration "(m~s^{-2})), 
             x = "Intertegular span (mm)") + 
        geom_ribbon(aes(ymin = blo, ymax = bhi), alpha = 0.2, color = NA)+
        geom_line()+
        theme(plot.background = element_rect(fill = "transparent",colour = NA), 
            panel.background = element_rect(fill = "transparent",colour = NA), 
            legend.position =c(0.2, 0.82), 
            legend.background = element_rect(fill="transparent", colour = NA),
            strip.background = element_blank(),
            legend.title = element_text(size = 8),
            panel.spacing.x = unit(1, "lines"), 
           plot.margin = margin(c(0.2,0.51, 0.2, 0.2), unit = "cm")) +
        scale_color_viridis_d(name = "Frequency range\nfor reward", end = 0.7, option = "inferno") + 
        scale_fill_viridis_d(name = "Frequency range\nfor reward", end =0.7, option = 'inferno')
ga1
```


# Compare acceleration analysis when 2 bees without IT are missing
```{r}
summary(m13)$coef
missing_accel <- update(m13, .~. - IT, data = sl[!is.na(sl$IT),])

summary(missing_accel)$coef
```


# Combine freq and accel plots into a single figure
```{r}
# set ggplot theme
theme_set(theme_classic() + 
            theme(axis.text=element_text(colour="black"), 
                  text=element_text(size=10)))

g2 <- g1 + 
  theme(legend.position = c(1.1,1.1), 
        legend.direction = "horizontal", 
        plot.margin = margin(c(1.5,0.51, 0.2, 0.2), unit = "cm"), 
        legend.title = element_text(size = 10), 
        legend.text = element_text(size = 8))

g2
ga2 <- ga1 + 
  theme(legend.position = "none", 
        plot.margin = margin(c(1.5,0.51, 0.2, 0.2), unit = "cm"))

ga2
aa <- plot_grid(g2, ga2)

aa

tiff(file.path(figDir, "001_Exp1_AccAndFreq.tiff"), 
      width = 6.5, height = 4, units = "in", res = 500)
aa
dev.off()
```


# print session info
```{r}
sessionInfo()
```

