---
output: html_document
editor_options: 
  chunk_output_type: console
---


<pre>
## Callin Switzer
## Analysis 2 of bee intrumental learning hypothesis experiment
### Does the rate of pollen collection change over time?

</pre>


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
```

```{r, warning = FALSE, message=FALSE}
#install packages
ipak <- function(pkg){
     new.pkg <- pkg[!(pkg %in% installed.packages()[, "Package"])]
     if(length(new.pkg)) install.packages(new.pkg, dependencies = TRUE)
     sapply(pkg, require, character.only = TRUE)
}

packages <- c("tidyverse")
ipak(packages)

# set ggplot theme
theme_set(theme_classic() + 
            theme(axis.text=element_text(colour="black"), 
                  text=element_text(size=10)))

# set  directories
dataDir <- file.path(getwd(), "data")
figDir <- file.path(getwd(), "figures")

print(paste("last run ", Sys.time()))
print(R.version)
```

# Read in data and tidy

```{r, message=FALSE}
# read in data
sl <- read_csv(file.path(dataDir, '01_InstrumentalLearning_annotated.csv'))

# clean data
sl <- sl %>%
  dplyr::select(-notes) %>%
  
  # make sure all bee colors are lowercase
  mutate(beeCol =  tolower(beeCol)) %>%
  
  # remove trials that contain errors
  filter(!(beeCol == "whitepink" & trialNum == 3)) %>% 
  filter(!(beeCol == "limepurple" & trialNum == 3)) %>% 
  filter(!(beeCol == "orange" & hive == 5)) %>%
  
  # color + hive is an ID var 
  mutate(beeColHive = interaction(beeCol, hive), 
         colNum = paste(beeCol, hive, sep = "_")) %>%
  mutate(beeColHive = factor(as.character(.$beeColHive), 
                             levels = sort(unique(as.character(.$beeColHive))))) %>%
  
  # fix IT for orange.3
  mutate(IT = replace(IT, beeColHive == "orange.3", 3.99)) %>%
  
  # convert hive to factor
  mutate(hive = as.factor(hive),
         trt2  = ifelse(trt == 'full' & trialNum > 1, 
                        "full_2", as.character(trt)), 
         trt = relevel(factor(trt), ref = "full")) %>%
  
  # divide amplitud by 2
  mutate(amp_acc2 = amp_acc / 2) %>%
  mutate(trialNum0 = trialNum - 1)



# calculate mean IT span (for centering)
IT_mean <- sl %>% 
    group_by(beeColHive) %>% 
    slice(1) %>%
  ungroup() %>%
  dplyr::select(IT_imputed) %>%
  summarise(mean = mean(IT_imputed)) %>%
  as.numeric()

# print overall mean of IT span
print(IT_mean)

# add new variable 
# that is the centered IT span
sl <- sl %>% 
  mutate(IT_centered = IT_imputed - IT_mean)

```


# Calculate more descriptive stats 
# and make supplementary fig for paper

```{r}
# std dev for rewards by treatment
s2 = sl %>%
  mutate(reward= grepl(pattern = "t", x = .$rewTF, ignore.case = TRUE)*1) 


s3 = s2 %>%
  group_by(beeColHive, trt2) %>%
  dplyr::summarize(mean_reward = mean(reward)) %>%
  group_by(trt2) %>%
  dplyr::arrange(desc(trt2)) 

# summary for paper
s3 %>%
  dplyr::summarize( mean_ = mean(mean_reward), 
                    reward_sd = sd(mean_reward), n = n()) %>%
  mutate(se = reward_sd / sqrt(n))


# total number of trials for each treatment
tapply((grepl("[tT]", as.character(sl$rewTF))), INDEX= sl$trt, length)

# total number of rewards per treatment
tapply((grepl("[tT]", as.character(sl$rewTF))), 
       INDEX= sl$trt, FUN = function(x) sum(x))

# total number of trials that were unrewarded per treatment
tapply((grepl("[tT]", as.character(sl$rewTF))), 
       INDEX= sl$trt, FUN = function(x) sum(!x))


# calculate avg freq & SE
se <- function(x) {
  sd(x) /  sqrt(length(x))
}


s4 = s2 %>%
  group_by(beeColHive, trt2) %>%
  dplyr::summarize(mean_freq = mean(freq), 
                   sd_freq = sd(freq)) %>%
  group_by(trt2) %>%
  dplyr::arrange(desc(trt2)) %>%
  filter(trt2 != "full") %>%
  ungroup() %>%
  mutate(trt3 = plyr::mapvalues(.$trt2, from = c("full_2", "high", "low"), 
                        to = c("Full range\n(220 - 450 Hz)", 
                               "High range\n(340 - 390 Hz)", 
                               "Low range\n(220 - 330 Hz)")))


# plot bee size in the groups
sl_sum <- sl %>%
  dplyr::select(beeColHive, trt2, IT_imputed) %>%
  distinct(beeColHive, trt2, .keep_all = TRUE) 

sl_sum %>%
  group_by(trt2) %>%
  summarize(counts = n())

sl %>% nrow

sl %>% group_by(beeColHive) %>%
  summarize(n()) %>%
  nrow()

sl %>% group_by(trt2) %>%
  summarize(count= n())

s2 <- sl_sum  %>%
  filter(trt2 != "full") %>%
  mutate(trt3 = plyr::mapvalues(.$trt2, from = c("full_2", "high", "low"), 
                        to = c("Full range\n(220 - 450 Hz)", 
                               "High range\n(340 - 390 Hz)", 
                               "Low range\n(220 - 330 Hz)")))
# join data  for plotting
s44 <- full_join(s2, s4)
```

# Plot rewards over time for each bee

```{r}
options(digits.secs=3)   

# calculate time differences
str_sub(sl$datetime, 21, 21) <- "."
sl <- sl %>%
  mutate(time_pos = as.POSIXct(datetime, format = "%Y_%m_%d__%H_%M_%OS" )) %>%
  group_by(beeColHive, trialNum) %>%
  arrange(beeColHive, trialNum, datetime) %>%
  mutate(time_diff = as.numeric(time_pos-lag(time_pos), units = 'secs')) %>%
  group_by(beeColHive) %>% 
  mutate(timeCumul = cumsum(replace_na(time_diff, replace = 0))) 

timeDiff = ggplot(sl[sl$rewTF, ], aes(x = timeCumul, y = time_diff, color = trt)) + 
  geom_point(alpha= 0.2, stroke = 0, aes( shape = trt )) + 
  facet_wrap(~beeColHive, ncol = 8, scales = "free_x") + 
  #ylim(c(0,5)) +
  #scale_x_continuous(breaks = seq(0, 10, 2)) +
  scale_color_viridis_d(name = "",
                        option = "magma", begin = 0.3,
                        end = 0.80,
                        guide = guide_legend(override.aes =
                                               list(alpha = 1, size = 4))) +
  theme_classic() +
  # theme(strip.background = element_blank(),
  #       strip.text = element_blank(),
  #       panel.border = element_rect(colour = "grey40",
  #                                   fill=NA, size=0.5),
  #       legend.position = 'top') +
  scale_shape_manual(name = "", values = c(16,17,15)) +
  labs(y = "Time between rewards (sec)", x = "Cumulative experience in experimental chamber (sec)") + 
  scale_y_log10()

timeDiff


# supplementary figure for paper
# ggsave(file.path(figDir, "freq_amp.tiff"), width = 6.5, 
#        height = 3, units = "in", dpi = 500)
```

## Plot rate of pollen collection

```{r}

sl <- sl %>%
  group_by(beeColHive) %>%
  arrange(beeColHive, datetime) %>%
  mutate(cumuPol =  cumsum(rewTF))

fta <- sl %>% 
  filter(trialNum == 1) %>% 
  dplyr::group_by(beeColHive) %>%
  dplyr::summarise(initialFreq = mean(freq)) %>%
  full_join(sl) %>%
  mutate(freqDiff = freq- initialFreq) %>%
  
  #keep bees that completed more than the first trial
  filter(beeColHive %in% .$beeColHive[.$trialNum > 1]) %>% 
  dplyr::arrange(trt) 


cc <- fta %>%
  ungroup() %>%
  filter(trialNum > 1) %>%
  group_by(beeColHive) %>%
  dplyr::summarize(meanFreq = mean(freq))

bb <- fta %>%
  dplyr::select(beeColHive, trt, trialNum, initialFreq) %>%
  dplyr::filter(trialNum == 2 | trialNum == 3) %>%
  group_by(beeColHive) %>%
  slice(1:1) %>%
  full_join(cc) %>%
  mutate(freqDiff = initialFreq - meanFreq) %>%
  dplyr::arrange(trt, desc(freqDiff))

fta <- fta %>%
  mutate(beeColHive = factor(beeColHive, levels = bb$beeColHive)) %>% 
  mutate(t5 = recode(.$trt, "full" = "Full range\n(220 - 450 Hz)", 
                       "high" = "High range\n(340 - 390 Hz)", 
                       "low" = "Low range\n(220 - 330 Hz)")) %>%
  filter(trialNum <= 10) %>%
  filter(!is.na(freq))

s44 = fta %>%
  group_by(trialNum, beeColHive, t5) %>%
  dplyr::summarise(meanFreq2 = mean(freq))
s44

pollenPlot = ggplot(fta[fta$rewTF,], aes(x = timeCumul / 60 , y = cumuPol, color = t5)) + 
  geom_step(size = 0.8) + 
  facet_wrap(~beeColHive, ncol = 8) + 
  scale_color_viridis_d(name = "",
                        option = "magma", begin = 0.3,
                        end = 0.80,
                        guide = guide_legend(override.aes =
                                               list(alpha = 1, size = 1.5))) +
  theme_classic() +
  theme(#strip.background = element_blank(),
    # text = element_text(size = 6), 
    # axis.title = element_text(size = 6), 
    # axis.text = element_text(size = 5),
    # legend.text = element_text(size = 6),
    strip.text = element_blank(),
    panel.border = element_rect(colour = "grey40",
                            fill=NA, size=0.5),
    axis.line = element_blank(),
    axis.ticks = element_line(size = 0.3),
  legend.position = 'top') +
  scale_shape_manual(name = "", values = c(16,17,15)) +
  labs(y = "Cumulative number of pollen rewards", x = "Cumulative time in experimental chamber (min)") + 
  scale_x_continuous(breaks = c(0, 25, 50, 75, 100), labels = c("", 25, 50, 75, ""))# 

pollenPlot

# supplementary figure for paper
ggsave(file.path(figDir, "CumulativePollen.tiff"), 
       width = 6.5*1.5,
       height = 4*1.5, units = "in", dpi = 500)

ggsave(file.path(figDir, "CumulativePollen.png"), 
       width = 6.5*1.5,
       height = 4*1.5, units = "in", dpi = 500)


timeDiff = ggplot(fta[fta$rewTF, ], aes(x = timeCumul / 60, y = time_diff, color = t5)) + 
  geom_point(alpha= 0.2, stroke = 0, aes( shape = t5 )) + 
  facet_wrap(~beeColHive, ncol = 8) + 
  scale_color_viridis_d(name = "",
                        option = "magma", begin = 0.3,
                        end = 0.80,
                        guide = guide_legend(override.aes =
                                               list(alpha = 1, size = 3))) +
  theme_classic() +
  theme(#strip.background = element_blank(),
    # text = element_text(size = 8), 
    # axis.title = element_text(size = 8), 
    # axis.text = element_text(size = 6),
    # strip.text = element_blank(),
    # legend.text = element_text(size = 8)
    panel.border = element_rect(colour = "grey40",
                            fill=NA, size=0.5),
    axis.line = element_blank(),
    axis.ticks = element_line(size = 0.3),
  legend.position = 'top') +
  scale_shape_manual(name = "", values = c(16,17,15)) +
  labs(y = "Time between rewards (sec)", x = "Cumulative experience in experimental chamber (min)") + 
  scale_y_log10() + 
  scale_x_continuous(breaks = c(0, 25, 50, 75, 100), labels = c("", 25, 50, 75, ""))# 

timeDiff

# supplementary figure for paper
ggsave(file.path(figDir, "TimeBWRewards.tiff"), plot = timeDiff,
       width = 6.5*1.5,
       height = 4*1.5, units = "in", dpi = 500)

ggsave(file.path(figDir, "TimeBWRewards.tiff"),  plot = timeDiff,
       width = 6.5*1.5,
       height = 4*1.5, units = "in", dpi = 500)

```


# Wright's Learning Curve
https://maaw.info/LearningCurveSummary.htm

```{r}
# Y = cumulative average time per reward
# X = cumulative number of units produced

wlc <- fta %>%
  group_by(beeColHive) %>%
  arrange(beeColHive, datetime) %>%
  mutate(attemptNum = seq(n())) %>%
  filter(rewTF) %>%
  group_by(beeColHive) %>%
  arrange(beeColHive, datetime) %>%
  mutate(rewNum2 = seq(n())) %>%
  mutate(cumulativeAvg = timeCumul / rewNum2) %>%
  mutate(cumAvgAttempts = attemptNum / rewNum2) %>%
  dplyr::select(beeColHive, datetime, time_diff, trialNum,
                timeCumul, rewTF, t5, rewNum2, cumulativeAvg,
                cumAvgAttempts, attemptNum)


PollenRate <-  ggplot(wlc, aes(x = rewNum2, y = attemptNum, color = t5)) + 
  geom_abline(intercept = 0, slope = 1, lty = 2, col = "grey", size = 0.2) +
  geom_point(alpha= 0.4, size = 0.8, stroke = 0, aes( shape = t5 )) + 
  facet_wrap(~beeColHive, ncol = 8) + 
  scale_color_viridis_d(name = "",
                        option = "magma", begin = 0.3 ,
                        end = 0.80,
                        guide = guide_legend(override.aes =
                                               list(alpha = 1, size = 4))) +
  theme_classic() +
  theme(strip.text = element_blank(),
    panel.border = element_rect(colour = "grey40",
                            fill=NA, size=0.5),
    axis.line = element_blank(),
    axis.ticks = element_line(size = 0.3),
    axis.text = element_text(color = "black"),
  legend.position = 'top') +
  labs(y = "Cumulative number of sonications", 
       x = "Cumulative number of rewards") + 
  scale_x_log10(limits = c(1, 3500)) + 
  scale_y_log10(limits = c(1, 3500)) + 
  scale_shape_manual(name = "", values = c(16, 17, 15))

 
PollenRate


# supplementary figure for paper
ggsave(file.path(figDir, "PollenRate.tiff"), plot = PollenRate,
       width = 6.5*1.5,
       height = 4*1.5, units = "in", dpi = 500)

ggsave(file.path(figDir, "PollenRate.png"),  plot = PollenRate,
       width = 6.5*1.5,
       height = 4*1.5, units = "in", dpi = 500)




wlc <- wlc %>% filter(cumulativeAvg != 0, 
         !is.na(time_diff))
```

```{r}

Wright_LC <-  ggplot(wlc, aes(y = cumulativeAvg, x = rewNum2, color = t5)) + 
  geom_line(alpha= 1, size = 0.8) + 
  facet_wrap(~beeColHive, ncol = 8) + 
  scale_color_viridis_d(name = "",
                        option = "magma", begin = 0.3,
                        end = 0.80,
                        guide = guide_legend(override.aes =
                                               list(alpha = 1, size = 3))) +
  theme_classic() +
  theme(strip.text = element_blank(),
    panel.border = element_rect(colour = "grey40",
                            fill=NA, size=0.5),
    axis.line = element_blank(),
    axis.ticks = element_line(size = 0.3),
  legend.position = 'top') +
  scale_y_log10() + 
  scale_x_log10() + 
  labs(y = "Cumulative time between pollen rewards", 
       x = "Cumulative number of rewards")

Wright_LC

```


# Calculations for Wright's learning curve
## based on attempt number (rather than time)

```{r}
# this method is correct for wright's learning curve
wlc2 <- fta %>%
  filter(trt != "full") %>%
  group_by(beeColHive) %>%
  arrange(beeColHive, datetime) %>%
  mutate(attemptNum = seq(n())) %>%
  filter(rewTF) %>%
  group_by(beeColHive) %>%
  arrange(beeColHive, datetime) %>%
  mutate(rewNum2 = seq(n())) %>%
  mutate(cumAvgAttempts = attemptNum / rewNum2) %>%
  dplyr::select(beeColHive, datetime, time_diff, trialNum,
                timeCumul, rewTF, t5, rewNum2, 
                cumAvgAttempts, attemptNum) 


# add extra levels so the plots align with other supp figs
wlc_extra = wlc2[1:2, ]
wlc_extra$cumAvgAttempts = 9999
wlc_extra$beeColHive = c("A", "B")
wlc3 <- rbind(wlc2, wlc_extra) %>% ungroup() %>%
  mutate(beeColHive = factor(beeColHive, levels = c("A", "B", levels(wlc$beeColHive))))


Wright_LC2 <-  ggplot(wlc3, aes(y = cumAvgAttempts, x = rewNum2, color = t5)) + 
  geom_line(alpha= 1, size = 0.8, aes(group = beeColHive)) + 
  facet_wrap(~beeColHive, ncol = 8) + 
  scale_color_viridis_d(name = "",
                        option = "magma", begin = (0.3 + 0.8)/2 ,
                        end = 0.80,
                        guide = guide_legend(override.aes =
                                               list(alpha = 1, size = 2))) +
  theme_classic() +
  theme(strip.text = element_blank(),
    panel.border = element_rect(colour = "grey40",
                            fill=NA, size=0.5),
    axis.line = element_blank(),
    axis.ticks = element_line(size = 0.3),
    axis.text = element_text(color = "black"),
  legend.position = 'top') +
  labs(y = "Cumulative average number of \nsonications between pollen rewards", 
       x = "Cumulative number of rewards") + 
  scale_y_log10(limits =  c(1, 50)) + 
  scale_x_log10() 
Wright_LC2


# supplementary figure for paper
ggsave(file.path(figDir, "LearningCurve_Wright.tiff"), plot = Wright_LC2,
       width = 6.5*1.5,
       height = 4*1.5*2/3, units = "in", dpi = 500)

ggsave(file.path(figDir, "LearningCurve_Wright.png"),  plot = Wright_LC2,
       width = 6.5*1.5,
       height = 4*1.5*2/3, units = "in", dpi = 500)


# plot on non-log scale
Wright_LC2_linearScale <-  ggplot(wlc3, aes(y = cumAvgAttempts, x = rewNum2, color = t5)) + 
  geom_line(alpha= 1, size = 0.8, aes(group = beeColHive)) + 
  facet_wrap(~beeColHive, ncol = 8) + 
  scale_color_viridis_d(name = "",
                        option = "magma", begin = (0.3 + 0.8)/2 ,
                        end = 0.80,
                        guide = guide_legend(override.aes =
                                               list(alpha = 1, size = 2))) +
  theme_classic() +
  theme(strip.text = element_blank(),
    panel.border = element_rect(colour = "grey40",
                            fill=NA, size=0.5),
    axis.line = element_blank(),
    axis.ticks = element_line(size = 0.3),
    axis.text = element_text(color = "black"),
  legend.position = 'top') +
  labs(y = "Cumulative average number of \nsonications between pollen rewards", 
       x = "Cumulative number of rewards")+ 
  ylim(c(0, 50))
Wright_LC2_linearScale


```


```{r, eval = FALSE}
# This method is wrong:
# this is how I calculated it the first time

wlc4 <- wlc3 %>% group_by(beeColHive) %>%
  arrange(datetime) %>%
  mutate(attemptDiff = as.numeric(attemptNum-lag(attemptNum))) %>%
  filter(!is.na(attemptDiff)) %>%
    mutate(CumulativeDiffs = cumsum(attemptDiff),
      cumAvgBuzzesBWRw = CumulativeDiffs / rewNum2)


Attempt_Diffs <-  ggplot(wlc4, aes(y = attemptDiff, x = rewNum2, color = t5)) + 
  geom_point(alpha= 1, size = 1, aes(group = beeColHive)) + 
  facet_wrap(~beeColHive, ncol = 8) + 
  scale_color_viridis_d(name = "",
                        option = "magma", begin = (0.3 + 0.8)/2 ,
                        end = 0.80,
                        guide = guide_legend(override.aes =
                                               list(alpha = 1, size = 3))) +
  theme_classic() +
  theme(strip.text = element_blank(),
    panel.border = element_rect(colour = "grey40",
                            fill=NA, size=0.5),
    axis.line = element_blank(),
    axis.ticks = element_line(size = 0.3),
    axis.text = element_text(color = "black"),
  legend.position = 'top') +
  labs(y = "Number of sonications between pollen rewards", 
       x = "Cumulative number of rewards")   + 
  ylim(c(0, 100))
Attempt_Diffs


Attempt_LC2 <-  ggplot(wlc4, aes(y = cumAvgBuzzesBWRw, x = rewNum2, color = t5)) + 
  geom_point(alpha= 0.3, size = 0.8, aes(group = beeColHive)) + 
  facet_wrap(~beeColHive, ncol = 8) + 
  scale_color_viridis_d(name = "",
                        option = "magma", begin = (0.3 + 0.8)/2 ,
                        end = 0.80,
                        guide = guide_legend(override.aes =
                                               list(alpha = 1, size = 3))) +
  theme_classic() +
  theme(strip.text = element_blank(),
    panel.border = element_rect(colour = "grey40",
                            fill=NA, size=0.5),
    axis.line = element_blank(),
    axis.ticks = element_line(size = 0.3),
    axis.text = element_text(color = "black"),
  legend.position = 'top') +
  labs(y = "Cumulative avg Number of sonications between pollen rewards", 
       x = "Cumulative number of rewards")  + 
  xlim(c(0, 64))
Attempt_LC2

```


# Make a plot with data from with website to see if it matches
# my calculations

```{r, eval = FALSE}
# what would an 80% Learning curve look like (from website)
# Y = cumulative average time per reward
# X = cumulative number of units produced
rewardNum <- c(1,2,3,4)
buzzesbwRewards <- c(100, 60, 50, 46)
cumulativeTime = cumsum(buzzesbwRewards)
cumAvgAttempts = cumulativeTime/rewardNum
plot(y = cumAvgAttempts, x = rewardNum, type = "b")
plot(log(cumAvgAttempts)~log(rewardNum), type = 'b')
lm(log(cumAvgAttempts) ~ log(rewardNum))
plot(lm(log(cumAvgAttempts) ~ log(rewardNum))$residuals)

# practice calculations to check work above
attemptNum = 1:sum(buzzesbwRewards)
rewTF = rep(0, length(attemptNum))
rewTF[cumsum(buzzesbwRewards)] = 1

fakeDF <- data.frame(attemptNum, rewTF)
fakeDF_2 <- fakeDF %>%
  mutate(attemptNum = seq(n())) %>%
  filter(rewTF == 1) %>%
  mutate(rewNum2 = seq(n())) %>%
  mutate(cumAvgAttempts = attemptNum / rewNum2) 

fakeDF_2

ggplot(fakeDF_2, aes(x = rewNum2, y = cumAvgAttempts)) + 
  geom_line() + 
  geom_point()


```


```{r}
# what would an 80% Learning curve look like (from website)
# Y = cumulative average time per reward
# X = cumulative number of units produced
rewardNum <- c(1,2,3,4,5,6,7,8,9,10)
buzzesbwRewards <- c(100, 200, 200, 200, 200, 
                     200, 200, 200, 200, 200)
cumulativeTime = cumsum(buzzesbwRewards)
cumAvgAttempts = cumulativeTime/rewardNum
plot(y = cumAvgAttempts, x = rewardNum, type = "b")
plot(log(cumAvgAttempts)~log(rewardNum), type = 'b')
lm(log(cumAvgAttempts) ~ log(rewardNum))
plot(lm(log(cumAvgAttempts) ~ log(rewardNum))$residuals)

# practice calculations to check work above
attemptNum = 1:sum(buzzesbwRewards)
rewTF = rep(0, length(attemptNum))
rewTF[cumsum(buzzesbwRewards)] = 1

fakeDF <- data.frame(attemptNum, rewTF)
fakeDF_2 <- fakeDF %>%
  mutate(attemptNum = seq(n())) %>%
  filter(rewTF == 1) %>%
  mutate(rewNum2 = seq(n())) %>%
  mutate(cumAvgAttempts = attemptNum / rewNum2) 

fakeDF_2

ggplot(fakeDF_2, aes(x = rewNum2, y = cumAvgAttempts)) + 
  geom_line() + 
  geom_point()


```


# print session info
```{r}
sessionInfo()
```

