---
output: html_document
editor_options: 
  chunk_output_type: console
---

<pre>
## Callin Switzer

## Data cleaning for 
## Analysis for experiment 3, physical constraint hypotheisis
### i.e. Does the mass of a flower affect bees sonication acceleration and frequency?

### Jan 17, 2019
</pre>


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
```


```{r, warning = FALSE, message=FALSE}
#install packages
ipak <- function(pkg){
     new.pkg <- pkg[!(pkg %in% installed.packages()[, "Package"])]
     if(length(new.pkg)) install.packages(new.pkg, dependencies = TRUE)
     sapply(pkg, require, character.only = TRUE)
}

packages <- c("ggplot2", 'lme4', 'sjPlot', 
              "multcomp", "plyr", "gamm4", 
              "cowplot", "tidyverse", 
              "dbscan")
ipak(packages)

# set ggplot theme
theme_set(theme_classic() + 
            theme(axis.text=element_text(colour="black"), 
                  text=element_text(size=10)))

# set  directories
dataDir <- file.path(getwd(), "data")
figDir <- file.path(getwd(), "figures")

print(paste("last run ", Sys.time()))
print(R.version)
```

# Read in data
# Calculate some descriptive statistics


```{r}
new_df = read.csv(file.path(dataDir, "02_1_HeavyLight_cleaned_posAdded.csv"))

# re-number colony number -- 3 & 4 should be 1 and 2
new_df$hive <- as.factor(as.numeric(as.character(new_df$hive)) - 2)

new_df$amp = new_df$MinMax_amp_Volts
new_df$amp_acc = new_df$MinMax_amp_acc_m.s.s

# divide amplitude by two
new_df$amp = new_df$amp / 2
new_df$amp_acc = new_df$amp_acc / 2

new_df$hive = as.factor(new_df$hive)
new_df$treatment <- mapvalues(new_df$treatment, from = c("sham", "weighted"), to = c("Sham", "Weighted"))

# print some descriptive statistics
nrow(new_df)
unique(new_df$hive)

# get sample size info
summdf <- new_df %>%
  group_by(beeID, treatment) %>%
  summarize(id_count = n())


sum(summdf$id_count)
length(unique(summdf$beeID))
nrow(summdf)

```

# Data engineering for analysis

```{r}
## add visitNumber counters for each bee and treatment
new_df <- new_df %>%
          arrange(beeID, datetime) %>%
          group_by(Folder, treatment) %>%
  # make visit number for each treatment
          mutate(visitNum_positive = seq(n())) %>%
          ungroup() 

# make visitNum_positive
newdf2 <- as.tbl(new_df)
 
newdf2$trt_char <- as.character(newdf2$treatment)
# get treatment ORDER
trts <- (sapply(unique(newdf2$beeID), FUN = function(x){
  startt = newdf2[newdf2$beeID == x & newdf2$rewNum == 1, "trt_char" ]
  return(as.character(startt))
}))

df4 <- data.frame(beeID = unique(newdf2$beeID), trts = trts)

newdf3 <- merge(newdf2, df4)

newdf3$trt2 <- as.factor(paste(newdf3$trts, "first"))

# calculate mean IT span (for centering)
IT_mean <- newdf3%>% 
    group_by(beeID) %>% 
    slice(1) %>%
  ungroup() %>%
  dplyr::select(IT) %>%
  summarise(mean = mean(IT)) %>%
  as.numeric()

print(IT_mean)

# add new variable 
# that is the centered IT span
newdf3 <- newdf3 %>% 
  mutate(IT_centered = IT - IT_mean)

```


```{r}
# make visitNum centered
newdf4 <- newdf3 %>%
          arrange(beeID, desc(datetime)) %>%
          group_by(Folder, treatment) %>%
  # make visit number for each treatment
          mutate(visitNum_neg = -seq(n())) %>%
          mutate(visitNum_centered= ifelse((treatment == "Weighted" & trt2 == "Weighted first") | 
                                             (treatment == "Sham" & trt2 == "Sham first"), yes = visitNum_neg, no =  visitNum_positive)) %>%
          ungroup() %>%
  arrange(beeID, datetime)
  


newdf4$treatment_I <- mapvalues(newdf3$treatment, from = c("Sham", "Weighted"), 
                          to = c("Sham", "Increased mass"))

newdf4$`Treatment Order` <- mapvalues(newdf3$trt2, from = c("Sham first", "Weighted first"), 
                          to = c("Sham first", "Increased mass first"))


timePlot_freq <- ggplot(newdf4, aes(x = visitNum_centered, y = freq)) + 
  geom_point(alpha = 0.3, position = position_jitter(width = 0, height = 4), 
             stroke = 0, size = 1.5, aes(color = treatment_I)) + 
  geom_line(stat="smooth",method = "loess",aes(color = treatment_I),
              size = 1.2, span = 0.9,
              alpha = 0.7, show.legend = FALSE)+
  facet_wrap(~beeID, labeller = label_both) + 
  theme(legend.position = "top", 
        panel.background = element_rect(fill = "transparent",colour = NA), 
        panel.border = element_rect(fill = "transparent"),
        legend.direction = "horizontal", 
        strip.text = element_blank()
        ) +  
  geom_vline(aes(xintercept = 0), lty = 2) + 
  xlim(c(-50, 50)) + 
  labs(y = expression ("Sonication frequency  "(Hz)), 
               x = "Visit Number, centered") + 
   scale_color_viridis_d(name = "Treatment", begin =0.3, end = 0.8, 
                        guide = guide_legend(override.aes = list(alpha = 1, size = 2)))

timePlot_freq


multiplier_Size = 1.5
ggsave(filename = file.path(figDir, "HeavyLight_IndividualPLots_raw.tiff"), width = 6.5 * multiplier_Size, height = 5* multiplier_Size)
ggsave(filename = file.path(figDir, "HeavyLight_IndividualPLots_raw.png"), width = 6.5* multiplier_Size, height = 5* multiplier_Size)

ggplot(newdf4, aes( freq, fill = treatment_I)) + 
  geom_histogram()+
  facet_wrap(~beeID, labeller = label_both) + 
  theme(legend.position = "top", 
        panel.background = element_rect(fill = "transparent",colour = NA), 
        legend.direction = "horizontal"
        ) +  
  geom_vline(aes(xintercept = 0), lty = 2) + 
  scale_color_viridis_d(name = "Treatment", begin =0.3, end = 0.8, 
                        guide = guide_legend(override.aes = list(alpha = 1, size = 2)))


# cluster with dbscan
bees <- unique(newdf4$beeID)
newdf4$cluster = 9999

for(ii in 1:length(bees)){
  beeID_ = bees[ii] 
  data_ = newdf4[newdf4$beeID == beeID_ , ]
  db_results = dbscan(as.matrix(data.frame(data_$freq)), eps = 29)
  
  newdf4$cluster[newdf4$beeID == beeID_ ] = db_results$cluster 
  
  lowClust = names(which.min(tapply(data_$freq, INDEX = db_results$cluster, mean))) %>% as.numeric
  # indicate low points or noise points if they're between the two groups
  newdf4$cluster[newdf4$beeID == beeID_ &  (newdf4$cluster == lowClust | newdf4$cluster == 0) ] = -1
  
  if(sd(db_results$cluster) == 0){
    newdf4$cluster[newdf4$beeID == beeID_ ] = 1
  }
  
  # data_2 = newdf4[newdf4$beeID == beeID_ & newdf4$treatment == "Sham", ]
  # db_results2 = dbscan(as.matrix(data_2$freq), eps = 20)
  # newdf4$cluster[newdf4$beeID == beeID_ & newdf4$treatment == "Sham"] = db_results2$cluster
  
}





newdf4 <- newdf4 %>%
  mutate(clustNames= recode(.$cluster, "-1" = "low_cluster", .default = "otherClust"))

# show points removed with dbscan
ggplot(newdf4, aes(x = visitNum_centered, y = freq, color = as.factor(clustNames))) + 
  geom_point(alpha = 1, position = position_jitter(width = 0, height = 1), 
             stroke = 0, size = 3) + 
  facet_wrap(~beeID, labeller = label_both) + 
  theme(legend.position = "top", 
        panel.background = element_rect(fill = "transparent",colour = NA), 
        legend.direction = "horizontal"
        ) +  
  geom_vline(aes(xintercept = 0), lty = 2) + 
  xlim(c(-50, 50)) + 
  labs(y = expression ("Sonication frequency  "(Hz)), 
               x = "Visit Number, centered") + 
   scale_color_viridis_d(name = "Cluster", begin =0.3, end = 0.8, 
                        guide = guide_legend(override.aes = list(alpha = 1, size = 2)))


## do gaussian mixture modeling to remove some more
ipak("mixtools")
set.seed(123)
newdf4$normCluster = 9999

for(beeID_ in c("14", "2", "29", "31", "white")){
    d3 <- newdf4[newdf4$beeID == beeID_ , ]
    mixmdl <- normalmixEM(d3$freq, k = 2)
    plot_mix_comps <- function(x, mu, sigma, lam) {
      lam * dnorm(x, mu, sigma)
    }
    
    smDF <- as.data.frame(mixmdl$posterior)
    low = smDF$comp.1 > 0.5
    
    data.frame(x = mixmdl$x) %>%
      ggplot() +
      geom_histogram(aes(x, ..density..), binwidth = 10, colour = "black", 
                     fill = "white") +
      stat_function(geom = "line", fun = plot_mix_comps,
                    args = list(mixmdl$mu[1], mixmdl$sigma[1], lam = mixmdl$lambda[1]),
                    colour = "red", lwd = 1.5) +
      stat_function(geom = "line", fun = plot_mix_comps,
                    args = list(mixmdl$mu[2], mixmdl$sigma[2], lam = mixmdl$lambda[2]),
                    colour = "blue", lwd = 1.5) +
      ylab("Density")
    
    newdf4[newdf4$beeID == beeID_ , "normCluster"] = low
  
  
}


# plot which points were removed by normal mixture model
ggplot(newdf4, aes(x = visitNum_centered, y = freq)) + 
  geom_point(alpha = 0.3, position = position_jitter(width = 0, height = 4), 
             stroke = 0, size = 4, aes(color = normCluster == 1 )) + 
  facet_wrap(~beeID, labeller = label_both) + 
  theme(legend.position = "top", 
        panel.background = element_rect(fill = "transparent",colour = NA), 
        legend.direction = "horizontal"
        ) +  
  geom_vline(aes(xintercept = 0), lty = 2) + 
  xlim(c(-50, 50)) + 
  labs(y = expression ("Sonication frequency  "(Hz)), 
               x = "Visit Number, centered") + 
   scale_color_viridis_d(name = "Treatment", begin =0.3, end = 0.8, 
                        guide = guide_legend(override.aes = list(alpha = 1, size = 2)))


# plot which points were removed (by both methods)
ggplot(newdf4, aes(x = visitNum_centered, y = freq)) + 
  geom_point(alpha = 0.3, position = position_jitter(width = 0, height = 4), 
             stroke = 0, size = 4, aes(color = normCluster == 1 | as.factor(clustNames) == "low_cluster" )) + 
  facet_wrap(~beeID, labeller = label_both) + 
  theme(legend.position = "top", 
        panel.background = element_rect(fill = "transparent",colour = NA), 
        legend.direction = "horizontal"
        ) +  
  geom_vline(aes(xintercept = 0), lty = 2) + 
  xlim(c(-50, 50)) + 
  labs(y = expression ("Sonication frequency  "(Hz)), 
               x = "Visit Number, centered") + 
   scale_color_viridis_d(name = "Point removed", begin =0.3, end = 0.8, 
                        guide = guide_legend(override.aes = list(alpha = 1, size = 2)))

ggplot(newdf4, aes(x = visitNum_centered, y = freq)) + 
  geom_point(alpha = 0.9, position = position_jitter(width = 0, height = 1), 
             stroke = 1, size = 3, aes(shape = treatment_I, 
                                       color =  normCluster == 1 | as.factor(clustNames) == "low_cluster" )) + 
  facet_wrap(~beeID, labeller = label_both) + 
  theme(legend.position = "top", 
        panel.background = element_rect(fill = "transparent",colour = NA), 
        panel.border = element_rect(fill = "transparent"),
        legend.direction = "horizontal", 
        strip.text = element_blank()
        ) +  
  geom_vline(aes(xintercept = 0), lty = 2) + 
  xlim(c(-50, 50)) + 
  labs(y = expression ("Sonication frequency  "(Hz)), 
               x = "Visit Number, centered") + 
   scale_color_viridis_d(name = "Point removed", begin =0.3, end = 0.8) + 
  scale_shape_manual(name = "Treatment", values = c(18, 20))


multiplier_Size = 1.5
ggsave(filename = file.path(figDir, "HeavyLight_IndividualPLots_Removed.tiff"), width = 6.5 * multiplier_Size, height = 5* multiplier_Size)
ggsave(filename = file.path(figDir, "HeavyLight_IndividualPLots_Removed.png"), width = 6.5* multiplier_Size, height = 5* multiplier_Size)


newdf5 <- newdf4[newdf4$cluster != -1, ]
newdf5 <- newdf5[newdf5$normCluster != 1, ]

# plot new data (cleaned)
ggplot(newdf5, aes(x = visitNum_centered, y = freq)) + 
  geom_point(alpha = 0.3, position = position_jitter(width = 0, height = 4), 
             stroke = 0, size = 4, aes(color = treatment_I)) + 
  geom_line(stat="smooth",method = "loess",aes(color = treatment_I),
              size = 1.2, span = 0.9,
              alpha = 0.7, show.legend = FALSE)+
  facet_wrap(~beeID, labeller = label_both) + 
  theme(legend.position = "top", 
        panel.background = element_rect(fill = "transparent",colour = NA), 
        legend.direction = "horizontal"
        ) +  
  geom_vline(aes(xintercept = 0), lty = 2) + 
  xlim(c(-50, 50)) + 
  labs(y = expression ("Sonication frequency  "(Hz)), 
               x = "Visit Number, centered") + 
   scale_color_viridis_d(name = "Treatment", begin =0.3, end = 0.8, 
                        guide = guide_legend(override.aes = list(alpha = 1, size = 2)))

ggplot(newdf5, aes(x = visitNum_centered, y = freq)) + 
  geom_point(alpha = 0.3, position = position_jitter(width = 0, height = 4), 
             stroke = 0, size = 4, aes(color = treatment_I)) + 
  geom_line(stat="smooth",method = "loess",aes(color = treatment_I),
              size = 1.2, span = 0.9,
              alpha = 0.7, show.legend = FALSE)+
  facet_wrap(~`Treatment Order`, labeller = label_both) + 
  theme(legend.position = "top", 
        panel.background = element_rect(fill = "transparent",colour = NA), 
        legend.direction = "horizontal"
        ) +  
  geom_vline(aes(xintercept = 0), lty = 2) + 
  xlim(c(-50, 50)) + 
  labs(y = expression ("Sonication frequency  "(Hz)), 
               x = "Visit Number, centered") + 
   scale_color_viridis_d(name = "Treatment", begin =0.3, end = 0.8, 
                        guide = guide_legend(override.aes = list(alpha = 1, size = 2)))


# make a new df3, and redo stats to see if there is a difference
newdf3_Original <- newdf3
newdf3 <- newdf5
newdf3$`Treatment Order` <- mapvalues(newdf3$trt2, from = c("Sham first", "Weighted first"), 
                          to = c("Sham first", "Increased mass first"))

nrow(newdf3_Original)
nrow(newdf3)

nrow(newdf3_Original) - nrow(newdf3)

```

# Modeling frequency with GAMM
# to look at non-linearities

```{r}
# visit number, treatment, IT span, colony, 
# the interaction of visit number and treatment order, 
# the interaction of treatment order and treatment, 
# and the interaction of IT span and treatment.

g00 = gamm4(freq ~
              trt2 + # treatment order
              treatment+   # treatment
              IT_centered + 
              hive + 
              s(visitNum_positive, by = interaction(trt2, treatment)) +
              trt2:treatment + 
              treatment:IT_centered,
              random =  ~(1|beeID), data = newdf3, 
              REML = FALSE)

par(mfrow = c(3,3))
aab <- plot(g00$gam, all.terms = TRUE, rug = FALSE, shade = TRUE)
summary(g00$gam) # Summary for paper 
summary(g00$mer)
dev.off()

# remove smooth interaction
g0 = gamm4(freq ~
              trt2 + # treatment order
              treatment +   # treatment
              IT_centered + 
              hive + 
              s(visitNum_positive, by =  treatment) +
              trt2:treatment + 
              treatment:IT_centered,
              random =  ~(1|beeID), data = newdf3, 
              REML = FALSE)

# start with gamm so I can show change by visit number
par(mfrow = c(3,3))
aab <- plot(g0$gam, all.terms = TRUE, rug = FALSE, shade = TRUE)
summary(g0$gam) # Summary for paper 
summary(g0$mer)
dev.off()

BIC(g00$mer, g0$mer) %>%
  mutate(model = row.names(.),
    keep = recode(as.character(order(.$BIC) == 1),
                   "TRUE" = "*",
                   "FALSE" = " "))# keep g0

g1 = gamm4(freq ~ 
               trt2 + # treatment order
               treatment + # treatment
               IT_centered + 
               hive + 
               s(visitNum_positive, by = trt2) +
               trt2:treatment + 
               treatment:IT_centered,
             random =  ~(1|beeID), data = newdf3, 
             REML = FALSE)

BIC(g0$mer, g1$mer) %>% 
  mutate(model = row.names(.), 
    keep = recode(as.character(order(.$BIC) == 1), 
                   "TRUE" = "*", 
                   "FALSE" = " "))# keep g0 -- smooth by treatment order
summary(g0$mer)$coef

g2 = gamm4(freq ~ 
               trt2 + # treatment order
               treatment + # treatment
               IT_centered + 
               hive + 
               s(visitNum_positive) + 
               trt2:treatment + 
               treatment:IT_centered,
             random =  ~(1|beeID), data = newdf3, 
             REML = FALSE)

par(mfrow = c(3,3))
aab <- plot(g2$gam, all.terms = TRUE, rug = FALSE, shade = TRUE)
dev.off()

BIC(g2$mer, g0$mer)%>% 
  mutate(model = row.names(.), 
    keep = recode(as.character(order(.$BIC) == 1), 
                   "TRUE" = "*", 
                   "FALSE" = " ")) # keep g2 -- no smooth by treatment order

g3 = gamm4(freq ~ 
               trt2 + # treatment order
               treatment + # treatment
               IT_centered + 
               as.factor(hive) + 
               visitNum_positive + 
               trt2:treatment + 
               treatment:IT_centered,
             random =  ~(1|beeID), data = newdf3, 
             REML = FALSE)

par(mfrow = c(3,3))
aab <- plot(g3$gam, all.terms = TRUE, rug = FALSE, shade = TRUE)
dev.off()

BIC(g2$mer, g3$mer)%>% 
  mutate(model = row.names(.), 
    keep = recode(as.character(order(.$BIC) == 1), 
                   "TRUE" = "*", 
                   "FALSE" = " ")) # keep g3 -- no smooth for visit number

```

# Modeling frequency with lmer

```{r modeling1}
# visit number, treatment, IT span, colony, 
# the interaction of visit number and treatment order, 
# the interaction of treatment order and treatment, 
# and the interaction of IT span and treatment.


# model with three-way interaction
m00 <- lmer(freq ~
              trt2 + # treatment order
              treatment + # treatment
              IT_centered + 
              hive + 
              visitNum_positive + 
              # two-way interactions
              treatment:IT_centered + 
              visitNum_positive:treatment +
              visitNum_positive:trt2 + 
              trt2:treatment + 
              # three way interaction
              trt2:treatment:visitNum_positive + 
              (1|beeID), data = newdf3, 
              REML = FALSE)

summary(m00)$coef

m0 = update(m00, .~. - trt2:treatment:visitNum_positive)
BIC(m0, m00) %>%
  mutate(model = row.names(.),
    keep = recode(as.character(order(.$BIC) == 1),
                   "TRUE" = "*",
                   "FALSE" = " ")) # keep m0 (no 3-way interaction)
summary(m0)$coef

m1 <- update(m0, .~. - trt2:visitNum_positive)
BIC(m1, m0) %>%
  mutate(model = row.names(.),
    keep = recode(as.character(order(.$BIC) == 1),
                   "TRUE" = "*",
                   "FALSE" = " ")) # keep m1
summary(m1)$coef

m2 <- update(m1, .~. - treatment:IT_centered)
BIC(m1, m2) %>%
  mutate(model = row.names(.),
    keep = recode(as.character(order(.$BIC) == 1),
                   "TRUE" = "*",
                   "FALSE" = " ")) # keep m2 
summary(m2)$coef

m3 <- update(m2, .~. - treatment:visitNum_positive)
BIC(m2, m3) %>%
  mutate(model = row.names(.),
    keep = recode(as.character(order(.$BIC) == 1),
                   "TRUE" = "*",
                   "FALSE" = " ")) # keep m3
summary(m3)$coef

m4 <- update(m3, .~. - trt2:treatment)
BIC(m3, m4) %>%
  mutate(model = row.names(.),
    keep = recode(as.character(order(.$BIC) == 1),
                   "TRUE" = "*",
                   "FALSE" = " ")) #keep m3
summary(m3)$coef

m5 <- update(m3, .~. - visitNum_positive)
BIC(m3, m5)%>%
  mutate(model = row.names(.),
    keep = recode(as.character(order(.$BIC) == 1),
                   "TRUE" = "*",
                   "FALSE" = " ")) # keep m5
summary(m5)$coef

m6 <- update(m5, .~. - IT_centered)
BIC(m5, m6) %>%
  mutate(model = row.names(.),
    keep = recode(as.character(order(.$BIC) == 1),
                   "TRUE" = "*",
                   "FALSE" = " ")) # keep m6
summary(m6)$coef

m7 <- update(m6, .~. - hive)
BIC(m7, m6) %>%
  mutate(model = row.names(.),
    keep = recode(as.character(order(.$BIC) == 1),
                   "TRUE" = "*",
                   "FALSE" = " ")) # keep m7
summary(m7)$coef

#______________________________________________
# Summarize frequency models for paper
#______________________________________________

# full frequency model
fullmod_freq = update(m00, .~., REML = TRUE)
summary(fullmod_freq)
round(summary(fullmod_freq)$coef, digits = 3)
# write.csv(data.frame(round(summary(fullmod_freq)$coef, digits = 3)), file = "coefs.csv")

                      
# small frequency model
finmod_freq <- update(m7, .~., REML = TRUE)
summary(finmod_freq) # final model for paper 

# frequency model that has the same predictors as accel model
matchingModel_freq =  lmer(freq ~ treatment + 
                             IT_centered  + 
                             treatment:IT_centered +  
                             trt2 + 
                             trt2:treatment + 
                              (1 | beeID), 
                          data = newdf3, REML = TRUE)


summary(matchingModel_freq)
# write.csv(data.frame(round(summary(matchingModel_freq)$coef, digits = 3)), file = "coefs.csv")

# 
summary(finmod_freq) 
plot(finmod_freq)
qqnorm(ranef(finmod_freq)$beeID[[1]])
qqline(ranef(finmod_freq)$beeID[[1]])

## check predictions
pframe_freq1 <- expand.grid(IT_centered  = 0,
                     treatment = levels(droplevels(newdf3$treatment)),
                     trt2 = levels(newdf3$trt2), 
                     beeID = 99999)

pframe_freq1$preds = predict(matchingModel_freq, newdata = pframe_freq1, re.form = NA)

pframe_freq1


# fit same model with original dataset to see differences
matchingModel_freq_originalData =  lmer(freq ~ treatment + 
                             IT_centered  + 
                             treatment:IT_centered +  
                             trt2 + 
                             trt2:treatment + 
                              (1 | beeID), 
                          data = newdf3_Original, REML = TRUE)

summary(matchingModel_freq_originalData)
# write.csv(data.frame(round(summary(matchingModel_freq_originalData)$coef, digits = 3)), file = "coefs.csv")

```

### Bootstrap CI's for figure for paper

```{r bootstrap_frequency, cache = TRUE}
# set number of bootstrap replicates for models
nbootSims = 10000
pframe_freq <- expand.grid(IT_centered  = seq(min(newdf3$IT_centered), 
                                         max(newdf3$IT_centered), 
                                         length.out = 50),
                     treatment = levels(droplevels(newdf3$treatment)),
                     trt2 = levels(newdf3$trt2), 
                     beeID = 99999)
pframe_freq$freq <- 0

# re.form sets all random effects to 0
pp <- predict(matchingModel_freq, 
                  newdata = pframe_freq,
                  re.form=NA, 
                  type = 'response')


### Calculate CI's (using bootstrap, not accounting for random effects)
bb2 <- bootMer(matchingModel_freq, 
               FUN=function(x) predict(x, 
                                       pframe_freq, 
                                       re.form=NA, 
                                       type = 'response'),
               nsim = nbootSims)
print(paste("Number of bootstrap samples", nrow(bb2$t)))
bb2_se <-apply(bb2$t,2,function(x) quantile(x, probs = c(0.025, 0.975)))
pframe_freq$blo<- (bb2_se[1,]) 
pframe_freq$bhi<- (bb2_se[2,])
pframe_freq$predMean <- pp
```

```{r}
# change "weighted" to "increased-mass"
pframe_freq$treatment <- mapvalues(pframe_freq$treatment, from = c("Sham", "Weighted"), 
                          to = c("Sham", "Increased mass"))

pframe_freq$`Treatment Order` = mapvalues(pframe_freq$trt2, 
                                           from = c("Sham first", "Weighted first"), 
                          to = c("Sham first", "Increased mass first"))
# plot freq vs. IT
# note: we add IT_mean to IT_centered to show IT on original scale
Freq_IT <- ggplot(pframe_freq, aes(x= IT_centered + IT_mean, y=predMean))+
              geom_ribbon(aes(x = IT_centered + IT_mean, ymin = blo, 
                              ymax = bhi, fill = treatment), alpha = 0.2) + 
              geom_line(aes(color = treatment), alpha = 1) + 
              labs(y =  expression ("Sonication frequency "(Hz)), x = "Intertegular span (mm)") + 
              scale_color_viridis_d(name = "Flower treatment", begin =0.3, end = 0.8) + 
              scale_fill_viridis_d(name = "Flower treatment", begin =0.3, end = 0.8) + 
              theme(legend.position = "top", 
                    legend.background = element_rect(fill=alpha('gray95', 1)), 
                    legend.direction="horizontal")  + 
  facet_wrap(~`Treatment Order`, labeller = label_both)
Freq_IT

# make plot with rug
weight_rug_freq <- Freq_IT + 
  geom_rug(data = newdf3[!duplicated(newdf3$IT),], 
           inherit.aes = FALSE, aes(x = IT))
weight_rug_freq  

Fig2a <- Freq_IT

# show figure
Fig2a
```

---

# Amplitude analysis

---


# Modeling amplitude with GAMM to look at non-linearities

```{r modeling-amp1}
# visit number, treatment, IT span, colony, 
# the interaction of visit number and treatment order, 
# the interaction of treatment order and treatment, 
# and the interaction of IT span and treatment.
# GAMM
g00 = gamm4(log(amp_acc)~
              trt2 + # treatment order
              treatment+   # treatment
              IT_centered + 
              hive + 
              s(visitNum_positive, by = interaction(trt2, treatment)) +
              trt2:treatment + 
              treatment:IT_centered,
              random =  ~(1|beeID), data = newdf3, 
              REML = FALSE)

par(mfrow = c(3,3))
aab <- plot(g00$gam, all.terms = TRUE, rug = FALSE, shade = TRUE)
summary(g00$gam) # Summary for paper 
summary(g00$mer)
dev.off()


# remove smooth interaction
g0 = gamm4(log(amp_acc) ~
              trt2 + # treatment order
              treatment +   # treatment
              IT_centered + 
              hive + 
              s(visitNum_positive, by =  treatment) +
              trt2:treatment + 
              treatment:IT_centered,
              random =  ~(1|beeID), data = newdf3, 
              REML = FALSE)

par(mfrow = c(3,3))
aab <- plot(g0$gam, all.terms = TRUE, rug = FALSE, shade = TRUE)
summary(g0$gam) 
summary(g0$mer)
dev.off()

BIC(g00$mer, g0$mer) %>%
  mutate(model = row.names(.),
    keep = recode(as.character(order(.$BIC) == 1),
                   "TRUE" = "*",
                   "FALSE" = " "))  # keep g0

g1 = gamm4(log(amp_acc) ~ 
               trt2 + # treatment order
               treatment + # treatment
               IT_centered + 
               hive + 
               s(visitNum_positive, by = trt2) +
               trt2:treatment + 
               treatment:IT_centered,
             random =  ~(1|beeID), data = newdf3, 
             REML = FALSE)

BIC(g0$mer, g1$mer) %>%
  mutate(model = row.names(.),
    keep = recode(as.character(order(.$BIC) == 1),
                   "TRUE" = "*",
                   "FALSE" = " "))  # keep g0 -- smooth by treatment order
summary(g0$mer)$coef

g2 = gamm4(log(amp_acc) ~ 
               trt2 + # treatment order
               treatment + # treatment
               IT_centered + 
               hive + 
               s(visitNum_positive) + 
               trt2:treatment + 
               treatment:IT_centered,
             random =  ~(1|beeID), data = newdf3, 
             REML = FALSE)

par(mfrow = c(3,3))
aab <- plot(g2$gam, all.terms = TRUE, rug = FALSE, shade = TRUE)
dev.off()

BIC(g2$mer, g0$mer) %>%
  mutate(model = row.names(.),
    keep = recode(as.character(order(.$BIC) == 1),
                   "TRUE" = "*",
                   "FALSE" = " "))  # keep g2 -- no smooth by treatment order

g3 = gamm4(log(amp_acc) ~ 
               trt2 + # treatment order
               treatment + # treatment
               IT_centered + 
               as.factor(hive) + 
               visitNum_positive + 
               trt2:treatment + 
               treatment:IT_centered,
             random =  ~(1|beeID), data = newdf3, 
             REML = FALSE)

par(mfrow = c(3,3))
aab <- plot(g3$gam, all.terms = TRUE, rug = FALSE, shade = TRUE)
dev.off()

BIC(g2$mer, g3$mer) %>%
  mutate(model = row.names(.),
    keep = recode(as.character(order(.$BIC) == 1),
                   "TRUE" = "*",
                   "FALSE" = " "))  # keep g3 -- no smooth for visit number

```

# Modeling amplitude with lmer

```{r modeling-amp3}
# visit number, treatment, IT span, colony, 
# the interaction of visit number and treatment order, 
# the interaction of treatment order and treatment, 
# and the interaction of IT span and treatment.

m00 <- lmer(log(amp_acc) ~
              trt2 + # treatment order
              treatment + # treatment
              IT_centered + 
              hive + 
              visitNum_positive + 
              # two-way interactions
              treatment:IT_centered + 
              visitNum_positive:treatment +
              visitNum_positive:trt2 + 
              trt2:treatment + 
              # three way interaction
              trt2:treatment:visitNum_positive + 
              (1|beeID), data = newdf3, 
              REML = FALSE)

summary(m00)$coef

ma0 = update(m00, .~. - trt2:treatment:visitNum_positive)
BIC(m00, ma0) %>%
  mutate(model = row.names(.),
    keep = recode(as.character(order(.$BIC) == 1),
                   "TRUE" = "*",
                   "FALSE" = " ")) # keep ma0 (no 3-way interaction)
summary(ma0)$coef

ma1 = update(ma0, .~. - trt2:visitNum_positive)
BIC(ma0, ma1) %>%
  mutate(model = row.names(.),
    keep = recode(as.character(order(.$BIC) == 1),
                   "TRUE" = "*",
                   "FALSE" = " "))  # keep ma1
summary(ma1)$coef

ma2 <- update(ma1, .~. - treatment:visitNum_positive)
BIC(ma1, ma2) %>%
  mutate(model = row.names(.),
    keep = recode(as.character(order(.$BIC) == 1),
                   "TRUE" = "*",
                   "FALSE" = " "))  # keep ma2
summary(ma2)$coef

ma3 <- update(ma2, .~. - trt2:treatment)
BIC(ma2, ma3) %>%
  mutate(model = row.names(.),
    keep = recode(as.character(order(.$BIC) == 1),
                   "TRUE" = "*",
                   "FALSE" = " "))  # keep ma3
summary(ma3)$coef

ma4 <- update(ma3, .~. - treatment:IT_centered)
BIC(ma3, ma4) %>%
  mutate(model = row.names(.),
    keep = recode(as.character(order(.$BIC) == 1),
                   "TRUE" = "*",
                   "FALSE" = " "))  # keep ma3
summary(ma3)$coef

ma5 <- update(ma3, .~. - visitNum_positive )
BIC(ma3, ma5) %>%
  mutate(model = row.names(.),
    keep = recode(as.character(order(.$BIC) == 1),
                   "TRUE" = "*",
                   "FALSE" = " "))  #keep ma5
summary(ma5)$coef

 
ma6 <- update(ma5, .~. - hive)
BIC(ma5, ma6) %>%
  mutate(model = row.names(.),
    keep = recode(as.character(order(.$BIC) == 1),
                   "TRUE" = "*",
                   "FALSE" = " ")) #keep ma6
summary(ma6)$coef

ma7 <- update(ma6, .~. - trt2)
BIC(ma7, ma6) %>%
  mutate(model = row.names(.),
    keep = recode(as.character(order(.$BIC) == 1),
                   "TRUE" = "*",
                   "FALSE" = " "))  #keep ma7
summary(ma7)$coef

#______________________________________________
# Summarize acceleration models for paper
#______________________________________________
fullMod_accel = update(m00, .~., REML = TRUE)
summary(fullMod_accel) # summary of full model for paper
# write.csv(data.frame(round(summary(fullMod_accel)$coef, digits = 3)), file = "coefs.csv")


finmod_accel_notFreq <- update(ma7, .~., REML = TRUE)
summary(finmod_accel_notFreq)
# write.csv(data.frame(round(summary(fullMod_accel)$coef, digits = 3)), file = "coefs.csv")



# diagnostics
plot(finmod_accel_notFreq)
qqnorm(ranef(finmod_accel_notFreq)$beeID[[1]])
qqline(ranef(finmod_accel_notFreq)$beeID[[1]])


# make model with same predictors as freq model
finmod_accel <- lmer(log(amp_acc) ~ treatment + 
                             IT_centered  + 
                             treatment:IT_centered +  
                             trt2 + 
                             trt2:treatment + 
                              (1 | beeID), 
                          data = newdf3, REML = TRUE)
summary(finmod_accel)

# write.csv(data.frame(round(summary(finmod_accel)$coef, digits = 3)), file = "coefs.csv")

# fit reduced model on original dataset to see differences
finmod_accel_originalData <- lmer(log(amp_acc) ~ treatment + 
                             IT_centered  + 
                             treatment:IT_centered +  
                             trt2 + 
                             trt2:treatment + 
                              (1 | beeID), 
                          data = newdf3_Original, REML = TRUE)
summary(finmod_accel_originalData)
nrow(newdf3_Original) - nrow(newdf3)

# write.csv(data.frame(round(summary(finmod_accel_originalData)$coef, digits = 3)), file = "coefs.csv")

```


### Predict amplitudes with some specific values
```{r}

summary(finmod_accel)$coef
pframe_accel0 <- expand.grid(IT_centered  = c(-1, 0, 1),
                     treatment = levels(droplevels(newdf3$treatment)),
                     trt2 = levels(newdf3$trt2),
                     beeID = 99999)

# exponentiate to put on original scale
# re.form sets all random effects to 0
pframe_accel0$preds <- predict(finmod_accel, 
                  newdata = pframe_accel0,
                  re.form=NA, 
                  type = 'response')

pframe_accel0 %>% 
  mutate(exp_preds = exp(preds)) %>%
  filter(IT_centered == 0)
exp(-0.28384542) * 17.37514

pframe_accel0 %>% 
  mutate(exp_preds = exp(preds)) %>%
  filter(IT_centered == -1)
exp(-0.28384542 + -1*0.33293768) * 16.420227

pframe_accel0 %>% 
  mutate(exp_preds = exp(preds)) %>%
  filter(IT_centered == 1)
exp(-0.28384542 + 1*0.33293768) * 18.38559


```



### Bootstrap CI's for amplitude for figure for paper

```{r bootstrap_amp2, cache = TRUE}
pframe_accel <- expand.grid(IT_centered  = seq(min(newdf3$IT_centered), 
                                         max(newdf3$IT_centered), 
                                         length.out = 50),
                     treatment = levels(droplevels(newdf3$treatment)),
                     trt2 = levels(newdf3$trt2), 
                     beeID = 99999)
pframe_accel$amp_acc <- 0

# exponentiate to put on original scale
# re.form sets all random effects to 0
pp <- exp(predict(finmod_accel, 
                  newdata = pframe_accel,
                  re.form=NA, 
                  type = 'response')) 


### Calculate CI's (using bootstrap, not accounting for random effects)
bb2 <- bootMer(finmod_accel, 
               FUN=function(x) predict(x, 
                                       pframe_accel, 
                                       re.form=NA, 
                                       type = 'response'),
               nsim = nbootSims)
print(paste("Number of bootstrap samples", nrow(bb2$t)))
bb2_se <-apply(bb2$t,2,function(x) quantile(x, probs = c(0.025, 0.975)))
pframe_accel$blo<- exp(bb2_se[1,]) # exponentiate to put on original scale
pframe_accel$bhi<- exp(bb2_se[2,])
pframe_accel$predMean <- pp
```


```{r}
# change "weighted" to "increased-mass"
pframe_accel$treatment <- mapvalues(pframe_accel$treatment, from = c("Sham", "Weighted"), 
                          to = c("Sham", "Increased mass"))

pframe_accel$`Treatment Order` = mapvalues(pframe_accel$trt2, 
                                           from = c("Sham first", "Weighted first"), 
                          to = c("Sham first", "Increased mass first"))


# plot amplitude vs. IT
# note: we add IT_mean to IT_centered to show IT on original scale
g0IT <- ggplot(pframe_accel, aes(x= IT_centered + IT_mean, y=predMean))+
          geom_ribbon(aes(x = IT_centered + IT_mean, ymin = blo, 
                          ymax = bhi, fill = treatment), alpha = 0.2) + 
          geom_line(aes(color = treatment), alpha = 1) + 
          labs(y = expression ("Sonication acceleration "(m~s^{-2})), 
               x = "Intertegular span (mm)") +
          scale_color_viridis_d(name = "Flower treatment",begin =0.3, end = 0.8) + 
          scale_fill_viridis_d(name = "Flower treatment", begin =0.3, end = 0.8) + 
          theme(legend.position = "top", 
            legend.background = element_rect(fill=alpha('gray95', 1)), 
            legend.direction="horizontal") + 
  facet_wrap(~`Treatment Order`, labeller = label_both)
g0IT

# make plot with rug
weight_rug <- g0IT + 
  geom_rug(data = newdf3[!duplicated(newdf3$IT),], 
           inherit.aes = FALSE, aes(x = IT))
  
Fig2b <- weight_rug

# show figure
Fig2b
```



### Put two plots into a single figure

```{r}
# set ggplot theme
theme_set(theme_classic() + 
            theme(axis.text=element_text(colour="black"), 
                  text=element_text(size=10)))

pfreq <- Fig2a + 
  theme(legend.position = c(0.83,0.8), 
        #plot.margin = margin(c(0.2,0.0, 0.2, 0.2), unit = "cm"), 
        legend.title = element_text(size = 10), 
        legend.text = element_text(size = 8), 
        plot.background = element_rect(fill = "transparent",colour = NA), 
        panel.background = element_rect(fill = "transparent",colour = NA), 
        legend.background = element_rect(fill="transparent", colour = NA),
        panel.spacing.x = unit(1, "lines"), 
        axis.text.x = element_blank(), 
        legend.direction = "vertical", 
        axis.title.x=element_blank(), 
        strip.text = element_text(size = 10)) 
  
pfreq
pamp <- Fig2b + 
        theme(legend.position = "none", 
              #plot.margin = margin(c(2,0.2, 0.2, 0.2), unit = "cm"), 
              plot.background = element_rect(fill = "transparent",colour = NA), 
              panel.background = element_rect(fill = "transparent",colour = NA), 
              legend.background = element_rect(fill="transparent", colour = NA),
              panel.spacing.x = unit(1, "lines"), 
              strip.background = element_blank(), 
              strip.text = element_blank())
              

  

aa <- plot_grid(pfreq, pamp,nrow = 2, align = 'v', axis = 'l', rel_widths = c(1,1))

ggdraw(aa)


save_plot(file.path(figDir, "003_HeavyLight_UPDATE.tiff"),
          plot = aa, base_width = 6.5, base_height = 5, dpi = 500)

save_plot(file.path(figDir, "003_HeavyLight_lowRes_UPDATE.png"),
          plot = aa, base_width = 6.5, base_height = 5, dpi = 500)

# 
# 
# save_plot(file.path(figDir, "003_HeavyLight_lowRes_big_UPDATE.png"), 
#           plot = aa, base_width = 8, base_height = 5, dpi = 500)
```


# Make plots of time vs. performance (raw data)
```{r}
# make visitNum centered

newdf4 <- newdf3 %>%
          arrange(beeID, desc(datetime)) %>%
          group_by(Folder, treatment) %>%
  # make visit number for each treatment
          mutate(visitNum_neg = -seq(n())) %>%
          mutate(visitNum_centered= ifelse((treatment == "Weighted" & trt2 == "Weighted first") | 
                                             (treatment == "Sham" & trt2 == "Sham first"), yes = visitNum_neg, no =  visitNum_positive)) %>%
          ungroup() %>%
  arrange(beeID, datetime)
  


newdf4$treatment_I <- mapvalues(newdf3$treatment, from = c("Sham", "Weighted"), 
                          to = c("Sham", "Increased mass"))

newdf4$`Treatment Order` <- mapvalues(newdf3$trt2, from = c("Sham first", "Weighted first"), 
                          to = c("Sham first", "Increased mass first"))


timePlot_freq <- ggplot(newdf4, aes(x = visitNum_centered, y = freq)) + 
  geom_point(alpha = 0.3, position = position_jitter(width = 0, height = 4), 
             stroke = 0, size = 1, aes(color = treatment_I)) + 
  geom_line(stat="smooth",method = "loess",aes(color = treatment_I),
              size = 1.2, span = 0.9,
              alpha = 0.7, show.legend = FALSE)+
  facet_wrap(~`Treatment Order`, labeller = label_both) + 
  theme(legend.position = "top", 
        panel.background = element_rect(fill = "transparent",colour = NA), 
        legend.direction = "horizontal"
        ) +  
  geom_vline(aes(xintercept = 0), lty = 2) + 
  xlim(c(-50, 50)) + 
  labs(y = expression ("Sonication frequency  "(Hz)), 
               x = "Visit Number, centered") + 
   scale_color_viridis_d(name = "Treatment", begin =0.3, end = 0.8, 
                        guide = guide_legend(override.aes = list(alpha = 1, size = 2)))
timePlot_freq

timePlot_merge = timePlot_freq + 
  theme(strip.background = element_rect(color = "black"), 
                               strip.text.x = element_text(margin = margin(0.2,0,0.2,0, "cm")), 
                               axis.text.x=element_blank(), 
                               axis.ticks.x=element_blank())



timePlot_acc <- ggplot(newdf4, aes(x = visitNum_centered, y = amp_acc)) + 
  geom_point(alpha = 0.3, stroke = 0, size = 1, aes(color = treatment_I)) + 
  geom_line(stat="smooth",method = "loess",aes(color = treatment_I),
                size = 1.2, span = 0.9,
                alpha = 0.7, show.legend = FALSE)+
  xlim(c(-50, 50)) + 
  scale_y_log10() + 
  facet_wrap(~`Treatment Order`, labeller = label_both) + 
  theme(panel.background = element_rect(fill = "transparent",colour = NA), 
        legend.position = "none", 
        legend.direction = "horizontal"
        ) + 
  geom_vline(aes(xintercept = 0), lty = 2) + 
  labs(y = expression ("Sonication acceleration  "(m~s^{-2})), 
               x = "Visit Number, centered") + 
  scale_color_viridis_d(name = "Treatment", begin =0.3, end = 0.8,
                        guide = guide_legend(override.aes = list(alpha = 1, size = 2)))

timePlot_acc



ab <- plot_grid(timePlot_merge + labs(x = ""), 
                timePlot_acc + theme(strip.background = element_blank(),
                                     strip.text.x = element_blank()), 
                align = 'h', axis = 'l', rel_widths = c(1,1), nrow = 2, 
                rel_heights = c(1.3, 1))

ggdraw(ab)


save_plot(file.path(figDir, "TimePlots_HeavyLight.tiff"), 
          plot = ab, base_width = 6.5, base_height = 5, dpi = 500)

save_plot(file.path(figDir, "TimePlots_HeavyLight_lowRes.png"), 
          plot = ab, base_width = 8, base_height = 5, dpi = 700)



```



```{r}
sessionInfo()
````
