---
output: html_document
editor_options: 
  chunk_output_type: console
---

<pre>
## Callin Switzer
## Analysis for experiment 3, physical constraint hypotheisis
### i.e. Does the mass of a flower affect bees sonication acceleration and frequency?

## Initial commit: 12 October 2017
## update 8 Dec 2017
## update 9 Aug 2018 -- make figures svg
## update 15 Dec 2018 -- make figures tiff
## Jan 8 update analysis to make it like other experiment
## Jan 14 2019 prep for zenodo
</pre>


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
```


```{r, warning = FALSE, message=FALSE}
#install packages
ipak <- function(pkg){
     new.pkg <- pkg[!(pkg %in% installed.packages()[, "Package"])]
     if(length(new.pkg)) install.packages(new.pkg, dependencies = TRUE)
     sapply(pkg, require, character.only = TRUE)
}

packages <- c("ggplot2", 'lme4', 'sjPlot', 
              "multcomp", "plyr", "gamm4", 
              "cowplot", "tidyverse")
ipak(packages)

# set ggplot theme
theme_set(theme_classic() + 
            theme(axis.text=element_text(colour="black"), 
                  text=element_text(size=10)))

# set  directories
dataDir <- file.path(getwd(), "data")
figDir <- file.path(getwd(), "figures")

print(paste("last run ", Sys.time()))
print(R.version)
```

# Read in data
# Calculate some descriptive statistics


```{r}
new_df = read.csv(file.path(dataDir, "02_1_HeavyLight_cleaned_posAdded.csv"))

# re-number colony number -- 3 & 4 should be 1 and 2
new_df$hive <- as.factor(as.numeric(as.character(new_df$hive)) - 2)

new_df$amp = new_df$MinMax_amp_Volts
new_df$amp_acc = new_df$MinMax_amp_acc_m.s.s

# divide amplitude by two
new_df$amp = new_df$amp / 2
new_df$amp_acc = new_df$amp_acc / 2

new_df$hive = as.factor(new_df$hive)
new_df$treatment <- mapvalues(new_df$treatment, from = c("sham", "weighted"), to = c("Sham", "Weighted"))

# print some descriptive statistics
nrow(new_df)
unique(new_df$hive)

# get sample size info
summdf <- new_df %>%
  group_by(beeID, treatment) %>%
  summarize(id_count = n())


sum(summdf$id_count)
length(unique(summdf$beeID))
nrow(summdf)

```

# Data engineering for analysis

```{r}
## add visitNumber counters for each bee and treatment
new_df <- new_df %>%
          arrange(beeID, datetime) %>%
          group_by(Folder, treatment) %>%
  # make visit number for each treatment
          mutate(visitNum_positive = seq(n())) %>%
          ungroup() 

# make visitNum_positive
newdf2 <- as.tbl(new_df)
 
newdf2$trt_char <- as.character(newdf2$treatment)
# get treatment ORDER
trts <- (sapply(unique(newdf2$beeID), FUN = function(x){
  startt = newdf2[newdf2$beeID == x & newdf2$rewNum == 1, "trt_char" ]
  return(as.character(startt))
}))

df4 <- data.frame(beeID = unique(newdf2$beeID), trts = trts)

newdf3 <- merge(newdf2, df4)

newdf3$trt2 <- as.factor(paste(newdf3$trts, "first"))

# calculate mean IT span (for centering)
IT_mean <- newdf3%>% 
    group_by(beeID) %>% 
    slice(1) %>%
  ungroup() %>%
  dplyr::select(IT) %>%
  summarise(mean = mean(IT)) %>%
  as.numeric()

print(IT_mean)

# add new variable 
# that is the centered IT span
newdf3 <- newdf3 %>% 
  mutate(IT_centered = IT - IT_mean)

```


# Modeling frequency with GAMM
# to look at non-linearities

```{r}
# visit number, treatment, IT span, colony, 
# the interaction of visit number and treatment order, 
# the interaction of treatment order and treatment, 
# and the interaction of IT span and treatment.

g00 = gamm4(freq ~
              trt2 + # treatment order
              treatment+   # treatment
              IT_centered + 
              hive + 
              s(visitNum_positive, by = interaction(trt2, treatment)) +
              trt2:treatment + 
              treatment:IT_centered,
              random =  ~(1|beeID), data = newdf3, 
              REML = FALSE)

par(mfrow = c(3,3))
aab <- plot(g00$gam, all.terms = TRUE, rug = FALSE, shade = TRUE)
summary(g00$gam) # Summary for paper 
summary(g00$mer)
dev.off()

# remove smooth interaction
g0 = gamm4(freq ~
              trt2 + # treatment order
              treatment +   # treatment
              IT_centered + 
              hive + 
              s(visitNum_positive, by =  treatment) +
              trt2:treatment + 
              treatment:IT_centered,
              random =  ~(1|beeID), data = newdf3, 
              REML = FALSE)

# start with gamm so I can show change by visit number
par(mfrow = c(3,3))
aab <- plot(g0$gam, all.terms = TRUE, rug = FALSE, shade = TRUE)
summary(g0$gam) # Summary for paper 
summary(g0$mer)
dev.off()

BIC(g00$mer, g0$mer) # keep g0

g1 = gamm4(freq ~ 
               trt2 + # treatment order
               treatment + # treatment
               IT_centered + 
               hive + 
               s(visitNum_positive, by = trt2) +
               trt2:treatment + 
               treatment:IT_centered,
             random =  ~(1|beeID), data = newdf3, 
             REML = FALSE)

BIC(g0$mer, g1$mer) # keep g0 -- smooth by treatment order
summary(g0$mer)$coef

g2 = gamm4(freq ~ 
               trt2 + # treatment order
               treatment + # treatment
               IT_centered + 
               hive + 
               s(visitNum_positive) + 
               trt2:treatment + 
               treatment:IT_centered,
             random =  ~(1|beeID), data = newdf3, 
             REML = FALSE)

par(mfrow = c(3,3))
aab <- plot(g2$gam, all.terms = TRUE, rug = FALSE, shade = TRUE)
dev.off()

BIC(g2$mer, g0$mer) # keep g2 -- no smooth by treatment order

g3 = gamm4(freq ~ 
               trt2 + # treatment order
               treatment + # treatment
               IT_centered + 
               as.factor(hive) + 
               visitNum_positive + 
               trt2:treatment + 
               treatment:IT_centered,
             random =  ~(1|beeID), data = newdf3, 
             REML = FALSE)

par(mfrow = c(3,3))
aab <- plot(g3$gam, all.terms = TRUE, rug = FALSE, shade = TRUE)
dev.off()

BIC(g2$mer, g3$mer) # keep g3 -- no smooth for visit number

```

# Modeling frequency with lmer

```{r modeling1}
# visit number, treatment, IT span, colony, 
# the interaction of visit number and treatment order, 
# the interaction of treatment order and treatment, 
# and the interaction of IT span and treatment.


# model with three-way interaction
m00 <- lmer(freq ~
              trt2 + # treatment order
              treatment + # treatment
              IT_centered + 
              hive + 
              visitNum_positive + 
              # two-way interactions
              treatment:IT_centered + 
              visitNum_positive:treatment +
              visitNum_positive:trt2 + 
              trt2:treatment + 
              # three way interaction
              trt2:treatment:visitNum_positive + 
              (1|beeID), data = newdf3, 
              REML = FALSE)

summary(m00)$coef

m0 = update(m00, .~. - trt2:treatment:visitNum_positive)
BIC(m0, m00) # keep m0 (no 3-way interaction)
summary(m0)$coef

m1 <- update(m0, .~. - trt2:visitNum_positive)
BIC(m1, m0) # keep m1
summary(m1)$coef

m2 <- update(m1, .~. - treatment:visitNum_positive)
BIC(m1, m2) # keep m2 
summary(m2)$coef

m3 <- update(m2, .~. - treatment:IT_centered)
BIC(m2, m3) # keep m3
summary(m3)$coef

m4 <- update(m3, .~. - trt2:treatment)
BIC(m3, m4) #keep m4
summary(m4)$coef

m5 <- update(m4, .~. - IT_centered)
BIC(m4, m5) # keep m5
summary(m5)$coef

m6 <- update(m5, .~. - hive)
BIC(m5, m6) # keep m6
summary(m6)$coef

m7 <- update(m6, .~. - trt2)
BIC(m7, m6) # keep m7
summary(m7)$coef

m8 <- update(m7, .~. - visitNum_positive)
BIC(m7, m8) # keep m8
summary(m8)$coef

m9 <- update(m8, .~. - treatment)
BIC(m9, m8) # keep m9
summary(m9)$coef

#______________________________________________
# Summarize frequency models for paper
#______________________________________________

# full frequency model
fullmod_freq = update(m00, .~., REML = TRUE)
summary(fullmod_freq)
round(summary(fullmod_freq)$coef, digits = 3)
# write.csv(data.frame(round(summary(fullmod_freq)$coef, digits = 3)), file = "coefs.csv")

                      
# small frequency model
finmod_freq <- update(m9, .~., REML = TRUE)
summary(finmod_freq) # final model for paper 

# frequency model that has the same predictors as accel model
matchingModel_freq =  lmer(freq ~ treatment + 
                             IT_centered  + 
                             treatment:IT_centered + 
                              (1 | beeID), 
                          data = newdf3, REML = TRUE)

summary(matchingModel_freq)


# no covariates predict frequency (when accounting for beeID)
summary(finmod_freq) 
plot(finmod_freq)
qqnorm(ranef(finmod_freq)$beeID[[1]])
qqline(ranef(finmod_freq)$beeID[[1]])
```

### Bootstrap CI's for figure for paper

```{r bootstrap_frequency, cache = TRUE}
# set number of bootstrap replicates for models
nbootSims = 10000
pframe_freq <- expand.grid(IT_centered  = seq(min(newdf3$IT_centered), 
                                         max(newdf3$IT_centered), 
                                         length.out = 50),
                     treatment = levels(droplevels(newdf3$treatment)),
                     beeID = 99999)
pframe_freq$freq <- 0

# re.form sets all random effects to 0
pp <- predict(matchingModel_freq, 
                  newdata = pframe_freq,
                  re.form=NA, 
                  type = 'response')


### Calculate CI's (using bootstrap, not accounting for random effects)
bb2 <- bootMer(matchingModel_freq, 
               FUN=function(x) predict(x, 
                                       pframe_freq, 
                                       re.form=NA, 
                                       type = 'response'),
               nsim = nbootSims)
print(paste("Number of bootstrap samples", nrow(bb2$t)))
bb2_se <-apply(bb2$t,2,function(x) quantile(x, probs = c(0.025, 0.975)))
pframe_freq$blo<- (bb2_se[1,]) 
pframe_freq$bhi<- (bb2_se[2,])
pframe_freq$predMean <- pp
```

```{r}
# change "weighted" to "increased-mass"
pframe_freq$treatment <- mapvalues(pframe_freq$treatment, from = c("Sham", "Weighted"), 
                          to = c("Sham", "Increased mass"))
# plot freq vs. IT
# note: we add IT_mean to IT_centered to show IT on original scale
Freq_IT <- ggplot(pframe_freq, aes(x= IT_centered + IT_mean, y=predMean))+
              geom_ribbon(aes(x = IT_centered + IT_mean, ymin = blo, 
                              ymax = bhi, fill = treatment), alpha = 0.2) + 
              geom_line(aes(color = treatment), alpha = 1) + 
              labs(y =  expression ("Sonication frequency "(Hz)), x = "Intertegular span (mm)") + 
              scale_color_viridis_d(name = "Flower treatment", begin =0.3, end = 0.8) + 
              scale_fill_viridis_d(name = "Flower treatment", begin =0.3, end = 0.8) + 
              theme(legend.position = "top", 
                    legend.background = element_rect(fill=alpha('gray95', 1)), 
                    legend.direction="horizontal") 
Freq_IT

# make plot with rug
weight_rug_freq <- Freq_IT + 
  geom_rug(data = newdf3[!duplicated(newdf3$IT),], 
           inherit.aes = FALSE, aes(x = IT))
  
Fig2a <- weight_rug_freq

# show figure
Fig2a
```

---

# Amplitude analysis

---


# Modeling amplitude with GAMM to look at non-linearities

```{r modeling-amp1}
# visit number, treatment, IT span, colony, 
# the interaction of visit number and treatment order, 
# the interaction of treatment order and treatment, 
# and the interaction of IT span and treatment.
# GAMM
g00 = gamm4(log(amp_acc)~
              trt2 + # treatment order
              treatment+   # treatment
              IT_centered + 
              hive + 
              s(visitNum_positive, by = interaction(trt2, treatment)) +
              trt2:treatment + 
              treatment:IT_centered,
              random =  ~(1|beeID), data = newdf3, 
              REML = FALSE)

par(mfrow = c(3,3))
aab <- plot(g00$gam, all.terms = TRUE, rug = FALSE, shade = TRUE)
summary(g00$gam) # Summary for paper 
summary(g00$mer)
dev.off()


# remove smooth interaction
g0 = gamm4(log(amp_acc) ~
              trt2 + # treatment order
              treatment +   # treatment
              IT_centered + 
              hive + 
              s(visitNum_positive, by =  treatment) +
              trt2:treatment + 
              treatment:IT_centered,
              random =  ~(1|beeID), data = newdf3, 
              REML = FALSE)

par(mfrow = c(3,3))
aab <- plot(g0$gam, all.terms = TRUE, rug = FALSE, shade = TRUE)
summary(g0$gam) 
summary(g0$mer)
dev.off()

BIC(g00$mer, g0$mer) # keep g0

g1 = gamm4(log(amp_acc) ~ 
               trt2 + # treatment order
               treatment + # treatment
               IT_centered + 
               hive + 
               s(visitNum_positive, by = trt2) +
               trt2:treatment + 
               treatment:IT_centered,
             random =  ~(1|beeID), data = newdf3, 
             REML = FALSE)

BIC(g0$mer, g1$mer) # keep g0 -- smooth by treatment order
summary(g0$mer)$coef

g2 = gamm4(log(amp_acc) ~ 
               trt2 + # treatment order
               treatment + # treatment
               IT_centered + 
               hive + 
               s(visitNum_positive) + 
               trt2:treatment + 
               treatment:IT_centered,
             random =  ~(1|beeID), data = newdf3, 
             REML = FALSE)

par(mfrow = c(3,3))
aab <- plot(g2$gam, all.terms = TRUE, rug = FALSE, shade = TRUE)
dev.off()

BIC(g2$mer, g0$mer) # keep g2 -- no smooth by treatment order

g3 = gamm4(log(amp_acc) ~ 
               trt2 + # treatment order
               treatment + # treatment
               IT_centered + 
               as.factor(hive) + 
               visitNum_positive + 
               trt2:treatment + 
               treatment:IT_centered,
             random =  ~(1|beeID), data = newdf3, 
             REML = FALSE)

par(mfrow = c(3,3))
aab <- plot(g3$gam, all.terms = TRUE, rug = FALSE, shade = TRUE)
dev.off()

BIC(g2$mer, g3$mer) # keep g3 -- no smooth for visit number

```

# Modeling amplitude with lmer

```{r modeling-amp3}
# visit number, treatment, IT span, colony, 
# the interaction of visit number and treatment order, 
# the interaction of treatment order and treatment, 
# and the interaction of IT span and treatment.

m00 <- lmer(log(amp_acc) ~
              trt2 + # treatment order
              treatment + # treatment
              IT_centered + 
              hive + 
              visitNum_positive + 
              # two-way interactions
              treatment:IT_centered + 
              visitNum_positive:treatment +
              visitNum_positive:trt2 + 
              trt2:treatment + 
              # three way interaction
              trt2:treatment:visitNum_positive + 
              (1|beeID), data = newdf3, 
              REML = FALSE)

summary(m00)$coef

ma0 = update(m00, .~. - trt2:treatment:visitNum_positive)
BIC(m00, ma0) # keep ma0 (no 3-way interaction)
summary(ma0)$coef

ma1 = update(ma0, .~. - trt2:visitNum_positive)
BIC(ma0, ma1) # keep ma1
summary(ma1)$coef

ma2 <- update(ma1, .~. - treatment:visitNum_positive)
BIC(ma1, ma2) # keep ma2
summary(ma2)$coef

ma3 <- update(ma2, .~. - trt2:treatment)
BIC(ma2, ma3) # keep ma3
summary(ma3)$coef

ma4 <- update(ma3, .~. - treatment:IT_centered)
BIC(ma3, ma4) # keep ma3
summary(ma3)$coef

ma5 <- update(ma3, .~. - trt2 )
BIC(ma3, ma5) #keep ma5
summary(ma5)$coef

 
ma6 <- update(ma5, .~. - hive)
BIC(ma5, ma6) #keep ma5
summary(ma6)$coef

ma7 <- update(ma6, .~. -visitNum_positive)
BIC(ma7, ma6) #keep ma7
summary(ma7)$coef

#______________________________________________
# Summarize acceleration models for paper
#______________________________________________
fullMod_accel = update(m00, .~., REML = TRUE)
summary(fullMod_accel) # summary of full model for paper

finmod_accel <- update(ma7, .~., REML = TRUE)
summary(finmod_accel)
# write.csv(data.frame(round(summary(fullMod_accel)$coef, digits = 3)), file = "coefs.csv")



# diagnostics
plot(finmod_accel)
qqnorm(ranef(finmod_accel)$beeID[[1]])
qqline(ranef(finmod_accel)$beeID[[1]])
```


### Predict amplitudes with some specific values
```{r}

summary(finmod_accel)$coef
pframe_accel0 <- expand.grid(IT_centered  = c(-1, 0, 1),
                     treatment = levels(droplevels(newdf3$treatment)),
                     beeID = 99999)

# exponentiate to put on original scale
# re.form sets all random effects to 0
pframe_accel0$preds <- predict(finmod_accel, 
                  newdata = pframe_accel0,
                  re.form=NA, 
                  type = 'response')

pframe_accel0 %>% 
  mutate(exp_preds = exp(preds)) %>%
  filter(IT_centered == 0)
exp(-0.28384542) * 17.37514

pframe_accel0 %>% 
  mutate(exp_preds = exp(preds)) %>%
  filter(IT_centered == -1)
exp(-0.28384542 + -1*0.33293768) * 16.420227

pframe_accel0 %>% 
  mutate(exp_preds = exp(preds)) %>%
  filter(IT_centered == 1)
exp(-0.28384542 + 1*0.33293768) * 18.38559


```



### Bootstrap CI's for amplitude for figure for paper

```{r bootstrap_amp2, cache = TRUE}
pframe_accel <- expand.grid(IT_centered  = seq(min(newdf3$IT_centered), 
                                         max(newdf3$IT_centered), 
                                         length.out = 50),
                     treatment = levels(droplevels(newdf3$treatment)),
                     beeID = 99999)
pframe_accel$amp_acc <- 0

# exponentiate to put on original scale
# re.form sets all random effects to 0
pp <- exp(predict(finmod_accel, 
                  newdata = pframe_accel,
                  re.form=NA, 
                  type = 'response')) 


### Calculate CI's (using bootstrap, not accounting for random effects)
bb2 <- bootMer(finmod_accel, 
               FUN=function(x) predict(x, 
                                       pframe_accel, 
                                       re.form=NA, 
                                       type = 'response'),
               nsim = nbootSims)
print(paste("Number of bootstrap samples", nrow(bb2$t)))
bb2_se <-apply(bb2$t,2,function(x) quantile(x, probs = c(0.025, 0.975)))
pframe_accel$blo<- exp(bb2_se[1,]) # exponentiate to put on original scale
pframe_accel$bhi<- exp(bb2_se[2,])
pframe_accel$predMean <- pp
```


```{r}
# change "weighted" to "increased-mass"
pframe_accel$treatment <- mapvalues(pframe_accel$treatment, from = c("Sham", "Weighted"), 
                          to = c("Sham", "Increased mass"))
# plot amplitude vs. IT
# note: we add IT_mean to IT_centered to show IT on original scale
g0IT <- ggplot(pframe_accel, aes(x= IT_centered + IT_mean, y=predMean))+
          geom_ribbon(aes(x = IT_centered + IT_mean, ymin = blo, 
                          ymax = bhi, fill = treatment), alpha = 0.2) + 
          geom_line(aes(color = treatment), alpha = 1) + 
          labs(y = expression ("Sonication acceleration "(m~s^{-2})), 
               x = "Intertegular span (mm)") +
          scale_color_viridis_d(name = "Flower treatment", begin =0.3, end = 0.8) + 
          scale_fill_viridis_d(name = "Flower treatment", begin =0.3, end = 0.8) + 
          theme(legend.position = "top", 
            legend.background = element_rect(fill=alpha('gray95', 1)), 
            legend.direction="horizontal") 
g0IT

# make plot with rug
weight_rug <- g0IT + 
  geom_rug(data = newdf3[!duplicated(newdf3$IT),], 
           inherit.aes = FALSE, aes(x = IT))
  
Fig2b <- weight_rug

# show figure
Fig2b
```



### Put two plots into a single figure

```{r}
# set ggplot theme
theme_set(theme_classic() + 
            theme(axis.text=element_text(colour="black"), 
                  text=element_text(size=10)))

pfreq <- Fig2a + 
  theme(legend.position = c(0.5,0.89), 
        legend.direction = "vertical", 
        #plot.margin = margin(c(0.2,0.0, 0.2, 0.2), unit = "cm"), 
        legend.title = element_text(size = 10), 
        legend.text = element_text(size = 8), 
        plot.background = element_rect(fill = "transparent",colour = NA), 
        panel.background = element_rect(fill = "transparent",colour = NA), 
        legend.background = element_rect(fill="transparent", colour = NA),
        strip.background = element_blank(),
        panel.spacing.x = unit(1, "lines")) +
  ylim(c(310, 390))

pamp <- Fig2b + 
        theme(legend.position = "none", 
              #plot.margin = margin(c(2,0.2, 0.2, 0.2), unit = "cm"), 
              plot.background = element_rect(fill = "transparent",colour = NA), 
              panel.background = element_rect(fill = "transparent",colour = NA), 
              legend.background = element_rect(fill="transparent", colour = NA),
              panel.spacing.x = unit(1, "lines"))

  

aa <- plot_grid(pfreq, pamp, align = 'h', axis = 'l', rel_widths = c(1,1))

ggdraw(aa)


save_plot(file.path(figDir, "003_HeavyLight.tiff"), 
          plot = aa, base_width = 6.5, base_height = 3, dpi = 500)

save_plot(file.path(figDir, "003_HeavyLight_lowRes.png"), 
          plot = aa, base_width = 6.5, base_height = 3, dpi = 150)
```


# Make plots of time vs. performance (raw data)
```{r}
# make visitNum centered

newdf4 <- newdf3 %>%
          arrange(beeID, desc(datetime)) %>%
          group_by(Folder, treatment) %>%
  # make visit number for each treatment
          mutate(visitNum_neg = -seq(n())) %>%
          mutate(visitNum_centered= ifelse((treatment == "Weighted" & trt2 == "Weighted first") | 
                                             (treatment == "Sham" & trt2 == "Sham first"), yes = visitNum_neg, no =  visitNum_positive)) %>%
          ungroup() %>%
  arrange(beeID, datetime)
  


newdf4$treatment_I <- mapvalues(newdf3$treatment, from = c("Sham", "Weighted"), 
                          to = c("Sham", "Increased mass"))

newdf4$`Treatment Order` <- mapvalues(newdf3$trt2, from = c("Sham first", "Weighted first"), 
                          to = c("Sham first", "Increased mass first"))


timePlot_freq <- ggplot(newdf4, aes(x = visitNum_centered, y = freq)) + 
  geom_point(alpha = 0.3, position = position_jitter(width = 0, height = 4), 
             stroke = 0, size = 1, aes(color = treatment_I)) + 
  geom_line(stat="smooth",method = "loess",aes(color = treatment_I),
              size = 1.2, span = 0.9,
              alpha = 0.7, show.legend = FALSE)+
  facet_wrap(~`Treatment Order`, labeller = label_both) + 
  theme(legend.position = "top", 
        panel.background = element_rect(fill = "transparent",colour = NA), 
        legend.direction = "horizontal"
        ) +  
  geom_vline(aes(xintercept = 0), lty = 2) + 
  xlim(c(-50, 50)) + 
  labs(y = expression ("Sonication frequency  "(Hz)), 
               x = "Visit Number, centered") + 
   scale_color_viridis_d(name = "Treatment", begin =0.3, end = 0.8, 
                        guide = guide_legend(override.aes = list(alpha = 1, size = 2)))
timePlot_freq

timePlot_merge = timePlot_freq + 
  theme(strip.background = element_rect(color = "black"), 
                               strip.text.x = element_text(margin = margin(0.2,0,0.2,0, "cm")), 
                               axis.text.x=element_blank(), 
                               axis.ticks.x=element_blank())



timePlot_acc <- ggplot(newdf4, aes(x = visitNum_centered, y = amp_acc)) + 
  geom_point(alpha = 0.3, stroke = 0, size = 1, aes(color = treatment_I)) + 
  geom_line(stat="smooth",method = "loess",aes(color = treatment_I),
                size = 1.2, span = 0.9,
                alpha = 0.7, show.legend = FALSE)+
  xlim(c(-50, 50)) + 
  scale_y_log10() + 
  facet_wrap(~`Treatment Order`, labeller = label_both) + 
  theme(panel.background = element_rect(fill = "transparent",colour = NA), 
        legend.position = "none", 
        legend.direction = "horizontal"
        ) + 
  geom_vline(aes(xintercept = 0), lty = 2) + 
  labs(y = expression ("Sonication acceleration  "(m~s^{-2})), 
               x = "Visit Number, centered") + 
  scale_color_viridis_d(name = "Treatment", begin =0.3, end = 0.8, 
                        guide = guide_legend(override.aes = list(alpha = 1, size = 2)))

timePlot_acc


ab <- plot_grid(timePlot_merge + labs(x = ""), 
                timePlot_acc + theme(strip.background = element_blank(),
                                     strip.text.x = element_blank()), 
                align = 'h', axis = 'l', rel_widths = c(1,1), nrow = 2, 
                rel_heights = c(1.3, 1))

ggdraw(ab)


save_plot(file.path(figDir, "TimePlots_HeavyLight.tiff"), 
          plot = ab, base_width = 6.5, base_height = 5, dpi = 500)

save_plot(file.path(figDir, "TimePlots_HeavyLight_lowRes.png"), 
          plot = ab, base_width = 6.5, base_height = 5, dpi = 200)



```



```{r}
sessionInfo()
````

